---
title: "IEVAR Numerical Tests"
author: "Andrew Roberts"
date: '2023-09-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(hetGP)

base_dir <- getwd() # Should be `pecan`
source(file.path(base_dir, "vsem_calibration_tests", "gp_emulator_functions.r"))
```

# Functions for EVAR computations. 

```{r}
calc_EVAR <- function(gpi, x_cond, x_int) {
  
  # Condition GP, using kriging believer approach. 
  pred_cond <- predict(gpi, x = x_cond)
  gpi_cond_KB <- update(gpi, Xnew = x_cond, Znew = pred_cond$mean, maxit = 0)
  
  # Predictions with kriging believer conditioned GP. 
  pred_KB_int <- predict(gpi_cond_KB, x = x_int)
  
  # Predict using current (unconditioned) GP. 
  pred_int <- predict(gpi, x = x_cond, xprime = x_int)
  
  # Compute EVAR. 
  inflation_factor <- 2 * drop(pred_int$cov)^2 / pred_int$sd2
  EVAR_vals <- transform_GP_predictions(pred_KB_int$mean, pred_KB_int$sd2, "LNP")$var * exp(inflation_factor)
  
  return(EVAR_vals)
  
}


calc_EVAR_MC <- function(gpi, x_cond, x_int, N_MC = 100000) {
  
  # Sample from current (unconditioned) GP at conditioning location. 
  pred_cond <- predict(gpi, x = x_cond)
  y_cond_samp <- rnorm(n = N_MC, pred_cond$mean, sqrt(pred_cond$sd2))
  
  # Condition GP on input conditioning location, leaving corresponding output NULL for now. 
  gpi_cond <- update(gpi, Xnew = x_cond, maxit = 0)
  gpi_cond$Z0 <- matrix(c(gpi_cond$Z0, NA), ncol = 1)
  gpi_cond$Z <- matrix(c(gpi_cond$Z, NA), ncol = 1)
  np1_idx <- nrow(gpi_cond$Z0)
  
  # Condition GP at each sample, compute conditioned LNP variance, and average over samples. 
  EVAR_approx_vals <- vector(mode = "numeric", length = nrow(x_int))
  for(i in seq_along(y_cond_samp)) {
    # Fill in output at conditioning input with current sample. 
    gpi_cond$Z0[np1_idx,1] <- y_cond_samp[i]
    gpi_cond$Z[np1_idx,1] <- y_cond_samp[i]
    
    # Compute predictive variance. 
    pred_samp_cond <- predict(gpi_cond, x = x_int)
    EVAR_samp <- transform_GP_predictions(pred_samp_cond$mean, pred_samp_cond$sd2, "LNP")$var
    EVAR_approx_vals <- EVAR_approx_vals + EVAR_samp/N_MC
  }
  
  return(EVAR_approx_vals)
  
}


calc_log_EVAR <- function(gpi, x_cond, x_int) {
  
  # Condition GP, using kriging believer approach. 
  pred_cond <- predict(gpi, x = x_cond)
  gpi_cond_KB <- update(gpi, Xnew = x_cond, Znew = pred_cond$mean, maxit = 0)
  
  # Predictions with kriging believer conditioned GP. 
  pred_KB_int <- predict(gpi_cond_KB, x = x_int)
  
  # Predict using current (unconditioned) GP. 
  pred_int <- predict(gpi, x = x_cond, xprime = x_int)
  
  # Compute EVAR. 
  inflation_factor <- 2 * drop(pred_int$cov)^2 / pred_int$sd2
  log_EVAR_vals <- convert_to_post_emulator_log_moments(pred_KB_int$mean, pred_KB_int$sd2, return_vals = "log_var")$log_var + inflation_factor
  
  return(log_EVAR_vals)
  
}


calc_log_EVAR_MC <- function(gpi, x_cond, x_int, N_MC = 100000) {
  
  # Sample from current (unconditioned) GP at conditioning location. 
  pred_cond <- predict(gpi, x = x_cond)
  y_cond_samp <- rnorm(n = N_MC, pred_cond$mean, sqrt(pred_cond$sd2))
  
  # Condition GP on input conditioning location, leaving corresponding output NULL for now. 
  gpi_cond <- update(gpi, Xnew = x_cond, maxit = 0)
  gpi_cond$Z0 <- matrix(c(gpi_cond$Z0, NA), ncol = 1)
  gpi_cond$Z <- matrix(c(gpi_cond$Z, NA), ncol = 1)
  np1_idx <- nrow(gpi_cond$Z0)
  
  # Condition GP at each sample, compute conditioned LNP variance, and average over samples. 
  log_EVAR_approx_vals <- matrix(nrow = N_MC, ncol = nrow(x_int))
  for(i in seq_along(y_cond_samp)) {
    # Fill in output at conditioning input with current sample. 
    gpi_cond$Z0[np1_idx,1] <- y_cond_samp[i]
    gpi_cond$Z[np1_idx,1] <- y_cond_samp[i]
    
    # Compute predictive variance. 
    pred_samp_cond <- predict(gpi_cond, x = x_int)
    log_EVAR_samp <- convert_to_post_emulator_log_moments(pred_samp_cond$mean, pred_samp_cond$sd2, return_vals = "log_var")$log_var
    log_EVAR_approx_vals[i,] <- log_EVAR_samp
  }
  
  log_EVAR_MC_estimates <- apply(log_EVAR_approx_vals, MARGIN = 2, FUN = matrixStats::logSumExp) - log(N_MC)
  
  return(log_EVAR_MC_estimates)
  
}

```



# Simple 1D Example

In this section we seek to calculate 
$$
\text{EVAR}(x, \tilde{x}) := \mathbb{E}_{\tilde{x} \sim f_n} \text{Var}\left[\exp\left\{f_n(x)\right\} | \tilde{x}, \tilde{y} \right]
$$

where $f_n \sim \mathcal{GP}(\mu_n, k_n)$ is a GP over some unknown function $f$ that has already been conditioned on an initial design, 
and $\tilde{y} := f(\tilde{x})$. EVAR admits the closed-form expression 
$$
\text{EVAR}(x, \tilde{x}) = \text{Var}\left[\exp\left\{f_n(x)\right\} | \tilde{x}, \mu_n(\tilde{x}) \right] \cdot \exp\left\{2 \frac{k_n^2(\tilde{x}, x)}{k_n(\tilde{x})}\right\}
$$

```{r}
# GP Setup. 

scale_factor <- 1.0
x_design <- matrix(c(0.0, 0.5, 1.0), ncol = 1)
x_grid <- matrix(seq(0, 1, length.out = 51), ncol = 1)

f_test <- function(x, scale = 1.0) scale * sin(2*x)

y_design <- f_test(x_design, scale = scale_factor)
y_grid <- f_test(x_grid, scale = scale_factor)

plot(x_grid, y_grid, type = "l", col = "red", xlab = "x", ylab = "y", main = "True Function and Design Points")
points(x_design, y_design, col = "black")

```

```{r}
# Fit GP
eps_nug <- sqrt(.Machine$double.eps)
gpi <- mleHomGP(x_design, y_design, covtype = "Gaussian", known = list(g = eps_nug))
gp_pred_grid <- predict(gpi, x = x_grid)

CI_upper <- qnorm(.99, gp_pred_grid$mean, sqrt(gp_pred_grid$sd2))
CI_lower <- qnorm(.01, gp_pred_grid$mean, sqrt(gp_pred_grid$sd2))  

plot(x_grid, y_grid, type = "l", col = "red", xlab = "x", ylab = "y", 
     main = "Initial GP Fit", ylim = c(min(CI_lower), max(CI_upper)))
points(x_design, y_design, col = "black")
lines(x_grid, gp_pred_grid$mean, col = "blue")
lines(x_grid, CI_upper, col = "gray")
lines(x_grid, CI_lower, col = "gray")

```

```{r}
# LNP Fit 
LNP_pred_list <- transform_GP_predictions(gp_pred_grid$mean, gp_pred_grid$sd2, "LNP") 

CI_upper_LNP <- qlnorm(.99, gp_pred_grid$mean, sqrt(gp_pred_grid$sd2))
CI_lower_LNP <- qlnorm(.01, gp_pred_grid$mean, sqrt(gp_pred_grid$sd2)) 

plot(x_grid, exp(y_grid), type = "l", col = "red", xlab = "x", ylab = "y", 
     main = "Initial LNP Fit", ylim = c(min(CI_lower_LNP), max(CI_upper_LNP)))
points(x_design, exp(y_design), col = "black")
lines(x_grid, LNP_pred_list$mean, col = "blue")
lines(x_grid, CI_upper_LNP, col = "gray")
lines(x_grid, CI_lower_LNP, col = "gray")

```


# Basic EVAR Plots. 


```{r}
test <- calc_EVAR_MC(gpi, x_cond = 0.75, x_int = x_grid, N_MC = 100000)

  plot(x_grid, LNP_pred_list$var, type = "l", col = "blue", xlab = "x", 
       ylab = "predictive variance (LNP)", main = "EVAR vs. Unconditioned Predictive Variance")
  lines(x_grid, test, col = "green")
  abline(v = 0.75, col = "green", lty = "dashed")

```


```{r}
x_cond_vals <- c(0.05, 0.3, 0.75, 0.9, 0.999)
IEVAR_vals <- vector(mode = "numeric", length = length(x_cond_vals))
IEVAR_vals_MC <- vector(mode = "numeric", length = length(x_cond_vals))

# EVAR plots. 
for(i in seq_along(x_cond_vals)) {
  x_cond <- x_cond_vals[i]
  EVAR_grid <- calc_EVAR(gpi, x_cond, x_grid)
  EVAR_grid_MC <- calc_EVAR_MC(gpi, x_cond, x_grid)
  IEVAR_vals[i] <- mean(EVAR_grid)
  IEVAR_vals_MC[i] <- mean(EVAR_grid_MC)
  
  plot(x_grid, LNP_pred_list$var, pch = 20, col = "blue", xlab = "x", 
       ylab = "predictive variance (LNP)", main = "EVAR vs. Unconditioned Predictive Variance")
  lines(x_grid, EVAR_grid, col = "green")
  lines(x_grid, EVAR_grid_MC, col = "purple")
  abline(v = x_cond, col = "green", lty = "dashed")
}

# IEVAR plot comparing conditioning points. 
plot(x_cond_vals, IEVAR_vals, col = "green", xlab = "x", ylab = "IEVAR", main = "IEVAR Evaluations")
points(x_cond_vals, IEVAR_vals_MC, col = "purple")

```


```{r}
# Same test but with log-scale computations. 

for(x_cond in x_cond_vals) {
  log_EVAR_grid <- calc_log_EVAR(gpi, x_cond, x_grid)
  log_EVAR_grid_MC <- calc_log_EVAR_MC(gpi, x_cond, x_grid)
  
  min_y <- min(min(log(LNP_pred_list$var)), min(log_EVAR_grid_MC), min(log_EVAR_grid))
  max_y <- max(max(log(LNP_pred_list$var)), max(log_EVAR_grid_MC), max(log_EVAR_grid))
  plot(x_grid, log(LNP_pred_list$var), pch = 20, col = "blue", xlab = "x", 
       ylab = "log predictive variance (LNP)", 
       main = "log(EVAR) vs. Log Unconditioned Predictive Variance", ylim = c(min_y, max_y))
  lines(x_grid, log_EVAR_grid, col = "green")
  lines(x_grid, log_EVAR_grid_MC, col = "purple")
  abline(v = x_cond, col = "green", lty = "dashed")
}

```

# Simple 1D Example with small or large numbers: makes log computations necessary. 

```{r}
# GP Setup. 

scale_factor <- 200
x_design <- matrix(c(0.0, 0.5, 1.0), ncol = 1)
x_grid <- matrix(seq(0, 1, length.out = 51), ncol = 1)

f_test <- function(x, scale = 1.0) scale * sin(2*x)

y_design <- f_test(x_design, scale = scale_factor)
y_grid <- f_test(x_grid, scale = scale_factor)

plot(x_grid, y_grid, type = "l", col = "red", xlab = "x", ylab = "y", main = "True Function and Design Points")
points(x_design, y_design, col = "black")

```

```{r}
# Fit GP
eps_nug <- sqrt(.Machine$double.eps)
gpi <- mleHomGP(x_design, y_design, covtype = "Gaussian", known = list(g = eps_nug))
gp_pred_grid <- predict(gpi, x = x_grid)
lnp_log_pred_grid <- convert_to_post_emulator_log_moments(gp_pred_grid$mean, gp_pred_grid$sd2)

CI_upper <- qnorm(.99, gp_pred_grid$mean, sqrt(gp_pred_grid$sd2))
CI_lower <- qnorm(.01, gp_pred_grid$mean, sqrt(gp_pred_grid$sd2))  

plot(x_grid, y_grid, type = "l", col = "red", xlab = "x", ylab = "y", 
     main = "Initial GP Fit", ylim = c(min(CI_lower), max(CI_upper)))
points(x_design, y_design, col = "black")
lines(x_grid, gp_pred_grid$mean, col = "blue")
lines(x_grid, CI_upper, col = "gray")
lines(x_grid, CI_lower, col = "gray")
```

```{r}
x_cond_vals <- c(0.05, 0.3, 0.75, 0.9, 0.999)
IEVAR_vals <- vector(mode = "numeric", length = length(x_cond_vals))
IEVAR_vals_MC <- vector(mode = "numeric", length = length(x_cond_vals))

for(x_cond in x_cond_vals) {
  log_EVAR_grid <- calc_log_EVAR(gpi, x_cond, x_grid)
  log_EVAR_grid_MC <- calc_log_EVAR_MC(gpi, x_cond, x_grid)
  
  min_y <- min(min(lnp_log_pred_grid$log_var), min(log_EVAR_grid_MC), min(log_EVAR_grid))
  max_y <- max(max(lnp_log_pred_grid$log_var), max(log_EVAR_grid_MC), max(log_EVAR_grid))
  plot(x_grid, lnp_log_pred_grid$log_var, pch = 20, col = "blue", xlab = "x", 
       ylab = "log predictive variance (LNP)", 
       main = "log(EVAR) vs. Log Unconditioned Predictive Variance", ylim = c(min_y, max_y))
  lines(x_grid, log_EVAR_grid, col = "green")
  lines(x_grid, log_EVAR_grid_MC, col = "purple")
  abline(v = x_cond, col = "green", lty = "dashed")
}

```
















```{r}
# TODO: 
# 1.) include estimation of other term in law of total variance, and check that the two sum to the total variance. 
# 2.) investigate the impact of using logs/logSumExp. 
# 3.) solution to issue with candidate points already being in design - simply exclude them. More interesting thing to look at 
#     is candidate points very close to existing design points. 
# 4.) add comparison of IEVAR estimates using logSumExp vs. not logSumExp. 
```













