---
title: "GP Accelerated MCMC Results"
author: "Andrew Roberts"
date: '2024-12-06'
output: html_document
params:
    args: 'vsem'
---

This document summarizes the outputs of the script `run_approx_mcmc`.r`, which 
runs various Gaussian process (GP) accelerated MCMC algorithms and saves the 
results from file. This script reads in these results, as well as other 
data previously saved, including: inverse problem object, log-likelihood
emulators, and exact MCMC samples from the true posterior for validation.

This document is intended to be run from 
the command line, as in:
`Rscript -e "rmarkdown::render('inv_prob_setup.Rmd.Rmd',params=list(args=<experiment_tag>))"`
where `<experiment_tag>` is replaced with the experiment tag string corresponding
with the specific outputs of `inv_prob_setup.Rmd`. The Rmarkdown output should 
be directed to 
`gp-calibration/output/gp_inv_prob/<experiment_tag>/init_emumator`. See 
[this](https://stackoverflow.com/questions/49904943/run-rmarkdown-with-arguments-on-the-command-line)
link for running Rmarkdown scripts from the command line and passing arguments.


```{r, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo=FALSE)

# TODO: make these comand line arguments.
design_tag <- file.path("init_emulator", "LHS_250")
run_id <- "test_run"
print(paste0("Design tag: ", design_tag))
print(paste0("MCMC Run ID: ", run_id))

# Settings: read from command line arguments.
args <-  params$args
experiment_tag <- args[1]
print(paste0("Experiment tag: ", experiment_tag))

# Specify burn-in for MCMC plots.
burn_in_start_exact <- 150000
```

```{r, echo = FALSE, include = FALSE}
library(ggplot2)
library(data.table)
library(assertthat)

# Setup and filepaths.
base_dir <- file.path("/projectnb", "dietzelab", "arober", "gp-calibration")
src_dir <- file.path(base_dir, "src")
experiment_dir <- file.path(base_dir, "output", "gp_inv_prob", experiment_tag)
setup_dir <- file.path(experiment_dir, "inv_prob_setup")
em_dir <- file.path(experiment_dir, "init_emulator")

# Source files.
source(file.path(src_dir, "general_helper_functions.r"))
source(file.path(src_dir, "inv_prob_test_functions.r"))
source(file.path(src_dir, "statistical_helper_functions.r"))
source(file.path(src_dir, "plotting_helper_functions.r"))
source(file.path(src_dir, "seq_design.r"))
source(file.path(src_dir, "gp_helper_functions.r"))
source(file.path(src_dir, "gpWrapper.r"))
source(file.path(src_dir, "llikEmulator.r"))
source(file.path(src_dir, "mcmc_helper_functions.r"))
source(file.path(src_dir, "gp_mcmc_functions.r"))
```


```{r, include=FALSE}
# Read inverse problem data. 
inv_prob <- readRDS(file.path(setup_dir, "inv_prob_list.rds"))
par_prior <- inv_prob$par_prior
llik_obj_exact <- inv_prob$llik_obj

# Read exact MCMC samples.
samp_dt_exact <- fread(file.path(setup_dir, "samp_exact.csv"))
samp_dt_exact <- select_mcmc_samp(samp_dt_exact, itr_start=burn_in_start_exact)

# Read GP-accelerated MCMC settings.
mcmc_approx_settings <- readRDS(file.path(experiment_dir, "mcmc_approx_settings.rds"))
```

# Single Design Results
Here we look at the approximate MCMC results from a single design realization.
It is infeasible to do this for each design realization individually, so 
future analysis only considers statistics computed from each MCMC run.

```{r}
# Identify all MCMC files.
design_id <- "1018157756"
mcmc_dir <- file.path(experiment_dir, run_id, design_tag, design_id)

mcmc_files <- list.files(mcmc_dir)
```

```{r}
# Set default burn-ins for each algorithm.
burn_in_start <- sapply(mcmc_approx_settings, function(x) as.integer(x$n_itr/2))
```

```{r}
# Summarize MCMC results, and compare to ground truth.

mcmc_file <- mcmc_files[1]
print(mcmc_file)

# Select samples.
samp_dt_approx <- readRDS(file.path(mcmc_dir, mcmc_file))$samp
samp_dt_approx <- select_mcmc_samp(samp_dt_approx, itr_start=burn_in_start)
print(samp_dt_approx[, .N, by=.(test_label, chain_idx)])
  
# MCMC diagnostics.
trace_plots <- get_trace_plots(samp_dt_approx)
  
# Comparison to exact MCMC.
samp_dt_approx <- combine_samp_dt(samp_dt_exact, samp_dt_approx)
kde1d_plots <- get_1d_kde_plots(samp_dt_approx, test_label_baseline="exact")
```

```{r}
for(plt in trace_plots) plot(plt)
for(plt in kde1d_plots) plot(plt)
```


```{r}
# TODO: should save following MCMC information from `run_approx_mcmc.r`:
# estimates means/vars/covs, R hat and ESS, coverage metrics, other posterior 
# error metrics.
#
# Should give a quick way to identify runs that produced weird results. In 
# particular, should compute means/covs by chain as an additional diagnostic 
# test for the MCMC sampling.
```




