---
title: "GP Accelerated MCMC Results"
author: "Andrew Roberts"
date: '2024-12-06'
output: html_document
---

While `approx_mcmc.Rmd` is intended to provide a high-level summary of the 
MCMC runs across all of the design replicates, this script is intended to 
provide an in-depth investigation of the results from individual designs. The 
emphasis is on gaining some understanding of the emulator fit and resulting 
approximate posterior geometry to better explain the behavior of the approximate
MCMC runs.

```{r, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo=FALSE)

# Settings
experiment_tag <- "vsem"
design_tag <- file.path("init_emulator", "LHS_250")
run_id <- "test_run7"
design_id <- "1018157756"

# Burn-in for exact MCMC samples.
itr_start <- c(exact=150000L)
```


```{r}
print(paste0("Experiment tag: ", experiment_tag))
print(paste0("Design tag: ", design_tag))
print(paste0("Design ID: ", design_id))
print(paste0("MCMC Run ID: ", run_id))
```
```{r, echo = FALSE, include = FALSE}
library(ggplot2)
library(data.table)
library(assertthat)

# Setup and filepaths.
base_dir <- file.path("/projectnb", "dietzelab", "arober", "gp-calibration")
src_dir <- file.path(base_dir, "src")
experiment_dir <- file.path(base_dir, "output", "gp_inv_prob", experiment_tag)
setup_dir <- file.path(experiment_dir, "inv_prob_setup")
em_dir <- file.path(experiment_dir, "init_emulator")
mcmc_dir <- file.path(experiment_dir, run_id, design_tag, design_id)

# Source files.
source(file.path(src_dir, "general_helper_functions.r"))
source(file.path(src_dir, "inv_prob_test_functions.r"))
source(file.path(src_dir, "statistical_helper_functions.r"))
source(file.path(src_dir, "plotting_helper_functions.r"))
source(file.path(src_dir, "seq_design.r"))
source(file.path(src_dir, "gp_helper_functions.r"))
source(file.path(src_dir, "gpWrapper.r"))
source(file.path(src_dir, "llikEmulator.r"))
source(file.path(src_dir, "mcmc_helper_functions.r"))
source(file.path(src_dir, "gp_mcmc_functions.r"))
```


```{r, include=FALSE}
# Read inverse problem data. 
inv_prob <- readRDS(file.path(setup_dir, "inv_prob_list.rds"))
par_prior <- inv_prob$par_prior
llik_obj_exact <- inv_prob$llik_obj
llik_exact <- llik_obj_exact$get_llik_func()
par_names <- inv_prob$par_names

# Read validation data.
test_info_prior <- readRDS(file.path(setup_dir, "test_info_prior.rds"))
test_info_post <- readRDS(file.path(setup_dir, "test_info_post.rds"))

# Read approximate llikEmulator object.
llik_em <- readRDS(file.path(experiment_dir, design_tag, design_id, "em_llik.rds"))

# Read llikEmulator design information.
design_info <- readRDS(file.path(experiment_dir, design_tag, 
                                 design_id, "design_info.rds"))

# Read llikEmulator prediction lists.
em_pred_list_post <- readRDS(file.path(experiment_dir, design_tag, design_id,
                                       "em_pred_list_post.rds"))
em_pred_list_prior <- readRDS(file.path(experiment_dir, design_tag, design_id,
                                       "em_pred_list_prior.rds"))

# Read llikEmulator validation results.
em_llik_errs_post <- readRDS(file.path(experiment_dir, design_tag, design_id,
                                       "em_llik_errs_post.rds"))
em_llik_errs_prior <- readRDS(file.path(experiment_dir, design_tag, design_id,
                                       "em_llik_errs_prior.rds"))

# Read exact MCMC samples.
samp_dt_exact <- fread(file.path(setup_dir, "samp_exact.csv"))
samp_dt_exact <- select_mcmc_samp(samp_dt_exact, itr_start=itr_start)

# Read GP-accelerated MCMC settings.
mcmc_approx_settings <- readRDS(file.path(experiment_dir, "mcmc_approx_settings.rds"))
```


```{r}
itr_start_approx <- sapply(mcmc_approx_settings, function(l) l$itr_start)

# Burn-ins for each algorithm.
print("----- Burn-ins -----")
print(paste0("Exact: ", itr_start))
print(itr_start_approx)
```


# Exact MCMC
```{r}
hist_plots_exact <- get_hist_plot_comparisons(samp_dt_exact, 
                                              test_label_baseline="prior")
for(plt in hist_plots_exact) plot(plt)
```

```{r}
summarize_approx_mcmc <- function(test_label, itr_start=1L, chain_idcs=NULL) {

  # Load and summarize MCMC samples from plug-in mean approximation.
  mcmc_path <- file.path(mcmc_dir, paste0("mcmc_samp_", test_label, ".rds"))

  # Read samples.
  mcmc_list <- readRDS(mcmc_path)
  samp_dt_approx <- select_mcmc_samp(mcmc_list$samp, itr_start=itr_start, 
                                     chain_idcs=chain_idcs)
  info_dt_approx <- select_mcmc_samp(mcmc_list$info, itr_start=itr_start, 
                                     chain_idcs=chain_idcs)
  mcmc_ic <- mcmc_list$output_list[[1]]$par_init
  print("-----> Initial conditions:")
  print(mcmc_ic)
  print("-----> Number of chains:")
  print(samp_dt_approx[, .N, by=.(test_label, chain_idx)])
  
  # Most algorithms return llik evaluations (or samples), but marginal 
  # acceptance prob chains only return acceptance probability evaluations.
  info_param_name <- "llik"
  if(test_label %in% c("marg-acc-prob-ind", "marg-acc-prob-joint")) {
    info_param_name <- "acc_prob"
  }
  
  # MCMC diagnostics.
  trace_plots <- get_trace_plots(samp_dt_approx)
  llik_trace_plot <- get_trace_plots(info_dt_approx, param_names=info_param_name)[[1]]
  prop_sd_trace_plots <- try(get_trace_plots(info_dt_approx, param_type="prop"))
  rhat_list <- calc_R_hat(samp_dt_approx, within_chain=TRUE)
  samp_dt_exact <- select_mcmc_samp(samp_dt_exact, test_labels="exact")
  samp_dt_approx <- combine_samp_dt(samp_dt_exact, samp_dt_approx)
  hist_plots <- get_hist_plot_comparisons(samp_dt_approx,
                                          test_label_baseline="exact")
  hist_plot_llik <- get_hist_plots(info_dt_approx, param_names=info_param_name,
                                   combine_chains=FALSE)[[1]]
  
  # Print R-hat statistics.
  print(rhat_list$R_hat_vals)
  
  # Display plots.
  for(plt in trace_plots) plot(plt)
  plot(llik_trace_plot)
  for(plt in hist_plots) plot(plt)
  plot(hist_plot_llik)
  for(plt in prop_sd_trace_plots) try(plot(plt))
  
  # Return data.
  return_list <- list(samp_dt_approx=samp_dt_approx, rhat_list=rhat_list)
  return(invisible(return_list))
}

```

# Plug-In Mean Posterior
We start by looking at the approximate posterior defined by plugging in the
GP mean, ignoring the GP variance.

```{r}
summarize_approx_mcmc("mean", itr_start=itr_start_approx)
```

# Marginal Posterior

```{r}
summarize_approx_mcmc("marginal", itr_start=itr_start_approx)
```

# Quantile Posterior

## 70th percentile
```{r}
summarize_approx_mcmc("quantile7", itr_start=itr_start_approx)
```

## 80th percentile
```{r}
summarize_approx_mcmc("quantile8", itr_start=itr_start_approx)
```

## 90th percentile
```{r}
summarize_approx_mcmc("quantile9", itr_start=itr_start_approx)
```

```{r}
# Scatter: true vs approximate log-likelihood.
# llik_em_mean_prior_plt <- llik_em$plot_pred_validation(input=test_info_prior$input,
#                                                        em_pred_list=em_pred_list_prior,
#                                                        true_llik=test_info_prior$llik, 
#                                                        include_interval=FALSE)
# llik_em_mean_post_plt <- llik_em$plot_pred_validation(input=test_info_post$input,
#                                                       em_pred_list=em_pred_list_post,
#                                                       true_llik=test_info_post$llik, 
#                                                       include_interval=FALSE)
```

# mcwmh-joint Posterior
```{r}
summarize_approx_mcmc("mcwmh-joint", itr_start=itr_start_approx)
```

# marg-acc-prob-joint Posterior
```{r}
summarize_approx_mcmc("marg-acc-prob-joint", itr_start=itr_start_approx)
```

# mcwmh-ind Posterior
```{r}
summarize_approx_mcmc("mcwmh-ind", itr_start=itr_start_approx)
```

# marg-acc-prob-ind Posterior
```{r}
summarize_approx_mcmc("marg-acc-prob-ind", itr_start=itr_start_approx)
```


# Pseudo-Marginal-independent Posterior

```{r}
summarize_approx_mcmc("pm-ind", itr_start=itr_start_approx)
```

# Pseudo-Marginal-joint Posterior

```{r}
summarize_approx_mcmc("pm-joint", itr_start=itr_start_approx, chain_idcs=2)
```





