---
title: "Inverse Problem Setup"
author: "Andrew Roberts"
date: '2024-11-22'
output: html_document
params:
    args: 'vsem'
---

This document summarizes the outputs of the script `inv_prob_setup.r`, which 
defines a Bayesian inverse problem and draws samples from its posterior 
using exact MCMC (no emulation). This document is intended to be run from 
the command line, as in:
`Rscript -e "rmarkdown::render('inv_prob_setup.Rmd.Rmd',params=list(args=<experiment_tag>))"`
where `<experiment_tag>` is replaced with the experiment tag string corresponding
with the specific outputs of `inv_prob_setup.Rmd`. The Rmarkdown output should 
be directed to 
`gp-calibration/output/gp_inv_prob/<experiment_tag>/inv_prob_setup`. See 
[this](https://stackoverflow.com/questions/49904943/run-rmarkdown-with-arguments-on-the-command-line)
link for running Rmarkdown scripts from the command line and passing arguments.

```{r, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo=FALSE)

# Settings: read from command line arguments.
args <-  params$args
experiment_tag <- args[1]
print(paste0("Experiment tag: ", experiment_tag))

# Specify burn-in for MCMC plots.
burn_in_start <- 150000
```

```{r, echo = FALSE, include = FALSE}
library(ggplot2)
library(data.table)
library(assertthat)

# Setup and filepaths.
base_dir <- file.path("/projectnb", "dietzelab", "arober", "gp-calibration")
src_dir <- file.path(base_dir, "src")
experiment_dir <- file.path(base_dir, "output", "gp_inv_prob", experiment_tag)
in_dir <- file.path(experiment_dir, "inv_prob_setup")

# Source files.
source(file.path(src_dir, "general_helper_functions.r"))
source(file.path(src_dir, "inv_prob_test_functions.r"))
source(file.path(src_dir, "statistical_helper_functions.r"))
source(file.path(src_dir, "plotting_helper_functions.r"))
source(file.path(src_dir, "seq_design.r"))
source(file.path(src_dir, "gp_helper_functions.r"))
source(file.path(src_dir, "gpWrapper.r"))
source(file.path(src_dir, "llikEmulator.r"))
source(file.path(src_dir, "mcmc_helper_functions.r"))
source(file.path(src_dir, "gp_mcmc_functions.r"))
```

# Summary of Inverse Problem 
```{r}
inv_prob <- readRDS(file.path(in_dir, "inv_prob_list.rds"))
par_prior <- inv_prob$par_prior
llik_obj_exact <- inv_prob$llik_obj
```

## Prior Distribution
```{r}
print(inv_prob$par_info)

prior_plots <- plot_prior_samp(par_prior)
for(plt in prior_plots) plot(plt)
```

## Likelihood
```{r}
print(llik_obj_exact$lik_description)
print(paste0("Likelihood variance ground truth: ", inv_prob$sig2_true))
print(paste0("Likelihood variance model: ", llik_obj_exact$get_lik_par()[1]))
```

## Model Driver
```{r}
driver_plt <- ggplot(data.frame(t=inv_prob$time_points,PAR=inv_prob$driver)) + 
                     geom_point(aes(x=t,y=PAR)) + 
                     ggtitle("Model Driver (PAR)") + 
                     xlab("Day") + ylab("PAR")

plot(driver_plt)
```

## Ground truth observable vs. observations
```{r}
df_inv_prob_data <- data.frame(t=inv_prob$time_points, 
                               y_true=drop(inv_prob$y_true), 
                               y=drop(inv_prob$y))
plt_outputs <- ggplot(data=df_inv_prob_data) +
                      geom_point(aes(x=t, y=y), color="red") + 
                      geom_line(aes(x=t, y=y_true), color="black") + 
                      ylab("output") + ggtitle("True and Observed y")

plot(plt_outputs)
```

## Other model outputs
```{r}
# Plot other model outputs. 
df_model_outputs_true <- data.frame(inv_prob$model_output_true[1,,])
colnames(df_model_outputs_true) <- inv_prob$output_names
df_model_outputs_true$time <- inv_prob$time_points
for(output_var in inv_prob$output_names) {
  output_var <- sym(output_var)
  plt <- ggplot(df_model_outputs_true) + geom_line(aes(x=time,y=!!output_var)) + 
         ggtitle("Ground Truth Trajectory") + xlab("Days") + ylab(output_var)
  plot(plt)
}

```

# Inverse Problem Solution via Exact MCMC (no emulation)

## Exact log-likelihood, varying one parameter at a time. 
```{r}
plt_proj <- llik_obj_exact$plot_1d_projection(n_points=100, 
                                              input_bounds=t(inv_prob$par_prior[,c("param1","param2")]), 
                                              plot_type="llik")
for(plt in plt_proj) plot(plt)
```

## Exact MCMC
```{r}
samp_dt_exact <- fread(file.path(in_dir, "samp_exact.csv"))

# MCMC Plots: exact. 
trace_plots_exact <- get_trace_plots(samp_dt_exact, test_labels="exact", 
                                     param_types="par",
                                     itr_start=burn_in_start)
hist_plots_exact <- get_hist_plot_comparisons(samp_dt_exact, test_labels="exact", 
                                              param_types="par", 
                                              itr_start=burn_in_start)
corr_plots_exact <- get_2d_density_plots(samp_dt_exact, test_labels="exact", 
                                         itr_start=burn_in_start)$exact

for(plt in trace_plots_exact) plot(plt) 
for(plt in hist_plots_exact) plot(plt)
for(plt in corr_plots_exact) plot(plt)
```

# Validation Design

## Design information
```{r}
test_info_prior <- readRDS(file.path(in_dir, "test_info_prior.rds"))
test_info_post <- readRDS(file.path(in_dir, "test_info_post.rds"))

print("Validation data sampled from prior:")
print(paste0("Number test points: ", test_info_prior$N_design))
print(paste0("Design method: ", test_info_prior$design_method))

print("Validation data sampled from posterior:")
print(paste0("Number test points: ", test_info_post$N_design))
print(paste0("Validation point sampling method: ", test_info_post$design_method))
```