---
title: "Basis Function Emulation"
author: "Andrew Roberts"
date: "2024-10-22"
output: html_document
---

This file provides an introduction to the functions defined in 
`basis_function_emulation.r`. The functions contained therein are designed 
to approximate the output space of functions with high-dimensional outputs
using a basis representation.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)

library(lhs)
library(ggplot2)
library(data.table)
library(assertthat)

base_dir <- file.path("/projectnb", "dietzelab", "arober", "gp-calibration")
src_dir <- file.path(base_dir, "src")

# Various helper functions.
source(file.path(src_dir, "seq_design.r"))
source(file.path(src_dir, "gpWrapper.r"))
source(file.path(src_dir, "llikEmulator.r"))
source(file.path(src_dir, "inv_prob_test_functions.r"))
source(file.path(src_dir, "general_helper_functions.r"))
source(file.path(src_dir, "plotting_helper_functions.r"))

# Main functions being illustrated.
source(file.path(src_dir, "basis_function_emulation.r"))
```

# Evaluate function at design points.
```{r}
# The function g() will be the test function used throughout this document.
inv_prob <- get_vsem_test_1()
g <- inv_prob$par_to_obs_op
```

```{r}
# Create an initial input design.
n_design <- 200L
U <- get_batch_design("LHS", N_batch=n_design, prior_params=inv_prob$par_prior)

# Run forward model g() at the design points.
G <- g(U)
```

```{r}
# Plot a subset of the function evaluations.
idx <- sample(1:nrow(G), size=20L)
plt_design <- plot_curves_1d_helper(inv_prob$time_points, t(G[idx,]), 
                                    plot_title="Model runs at design points",
                                    xlab="days", ylab="LAI")
              
plot(plt_design + theme(legend.position="none"))
```

# Run PCA on model outputs
```{r}
pca_list <- pca(G) 
sdev <- pca_list$sqrt_val # square root of eigenvalues of empirical covariance.
V <- pca_list$vec # eigenvectors.
```

```{r}
# Eigenvalue (scree) plot.
scree_plt <- ggplot(data.frame(k=seq_along(sdev), eigenval=sdev)) + 
             geom_point(aes(x=k, y=eigenval)) + ylab("Sqrt eigenvalues")

plot(scree_plt)
```

```{r}
# Plot the dominant eigenvectors.
r <- 3
plt_eig <- plot_curves_1d_helper(inv_prob$time_points, V[,1:r, drop=FALSE], 
                                 plot_title="Dominant eigenvectors",
                                 xlab="days", ylab="LAI")
plot(plt_eig)
```

We define a cutoff and only include the dominant principal components.
```{r}
# Basis vectors.
B <- V[,1:r, drop=FALSE]
sdev_r <- sdev[1:r]
```


Below we plot the basis weight functions $w_j(u)$ over a grid of $u$ values. 
Since these functions map from $\mathbb{R}^r$ to $\mathbb{R}$ we cannot easily 
plot them when $r > 1$. We therefore plot 1d projections by varying only one 
input variable at a time. 
```{r}
# First, get the grids of input points to plot. Note that the below code for 
# defining `par_bounds` only works for uniform priors.
par_bounds <- t(inv_prob$par_prior[,c("param1","param2")])
U_proj_list <- get_input_grid_1d_projection(inv_prob$par_names, X_bounds=par_bounds)

# Plot 1d projections of w_j(u), for j=1,...,r.
for(par in names(U_proj_list)) {
  U_proj <- U_proj_list[[par]][[1]]
  w_U <- eval_basis_weights(U_proj, g, B, m=pca_list$mean)
  plt <- plot_curves_1d_helper(U_proj[,par], w_U, df_by=NULL, 
                               plot_title=paste0("w(u), projection on ", par),
                               xlab="u", ylab=par)
  plot(plt)
}

```














