---
title: "eki_gp_emulation"
author: "Andrew Roberts"
date: "2025-02-08"
output: html_document
---

This tutorial demonstrates how ensemble Kalman inversion 
(EKI; implemented in `ens_Kalman_inversion.r`) and Gaussian process (GP)
emulation can be combined to approximate the posterior distribution in a 
Bayesian inverse problem. In particular, EKI is used to cheaply generate 
approximate samples from the posterior distribution. These samples are then 
used as an initial design to fit a GP emulator of the forward model or 
log-likelihood, which can then be used in MCMC algorithms.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)

library(data.table)
library(ggplot2)

# Filepath definitions.
base_dir <- file.path("/projectnb", "dietzelab", "arober", "gp-calibration")
src_dir <- file.path(base_dir, "src")

# Source required files.
source(file.path(src_dir, "general_helper_functions.r"))
source(file.path(src_dir, "inv_prob_test_functions.r"))
source(file.path(src_dir, "statistical_helper_functions.r"))
source(file.path(src_dir, "plotting_helper_functions.r"))
source(file.path(src_dir, "seq_design.r"))
source(file.path(src_dir, "gp_helper_functions.r"))
source(file.path(src_dir, "ens_Kalman_inversion.r"))
source(file.path(src_dir, "gpWrapper.r"))
source(file.path(src_dir, "llikEmulator.r"))
source(file.path(src_dir, "mcmc_helper_functions.r"))
source(file.path(src_dir, "gp_mcmc_functions.r"))
```

# Setup: load inverse problem
```{r}
# Load inverse problem setup information. Store log-likelihood upper bound.
inv_prob <- get_vsem_test_1()
par_names <- inv_prob$par_names
par_prior <- inv_prob$par_prior
llik_bounds <- inv_prob$llik_obj$get_llik_bounds()

print(paste0("Input dimension: ", inv_prob$dim_par))
print(paste0("Observation space dimension: ", inv_prob$dim_obs))
print(paste0("Log-likelihood bounds: ", paste(llik_bounds, collapse=", ")))
```

```{r}
# Create initial design.
n_design <- 250L
design_method <- "LHS"
design_info <- get_init_design_list(inv_prob, design_method, n_design)

print(paste0("Number design points: ", n_design))
print(paste0("Design method: ", design_method))
```

```{r}
# Sample test points from prior.
n_design_test <- 500L
test_info_prior <- get_init_design_list(inv_prob, design_method, n_design_test)
```

# Exact Posterior Samples
We start by generating samples using standard (exact) MCMC to be used as a 
baseline for comparison when evaluating the approximation approaches.

```{r}
# Run exact MCMC.
mcmc_settings <- list(test_label="exact", mcmc_func_name="mcmc_bt_wrapper", 
                      sampler="DEzs", n_itr=50000L, try_parallel=FALSE,  
                      n_chain=4L, defer_ic=TRUE)

# MCMC sampling using exact likelihood.
mcmc_list <- run_mcmc(inv_prob$llik_obj, par_prior, mcmc_settings)

# Table storing MCMC samples.
burn_in_exact <- round(mcmc_settings$n_itr/2)
samp_dt <- select_mcmc_itr(mcmc_list$samp, itr_start=burn_in_exact)
mcmc_metadata <- mcmc_list$output_list
```

```{r}
# Test points subsampled from true posterior.
samp_mat <- select_mcmc_samp_mat(samp_dt, test_label="exact", 
                                 param_type="par")[,par_names]
test_info_exact <- get_init_design_list(inv_prob, design_method="subsample", 
                                        N_design=n_design_test,
                                        design_candidates=samp_mat)
```

# Comparing EKI vs. GP approach: single round of forward model runs
We start by creating a "space-filling design" based on the prior distribution.
We use this as an initial ensemble in the EKI algorithm. We also fit a GP 
emulator, so that we can compare the EKI and GP approximations after a single 
round of forward model runs.

## GP emulator

### Fit GP
```{r}
# Fit GP for log-likelihood.
gp_obj <- gpWrapperKerGP(design_info$input, matrix(design_info$llik, ncol=1), 
                         scale_input=TRUE, normalize_output=TRUE)
gp_obj$set_gp_prior("Gaussian", "quadratic", include_noise=FALSE)
gp_obj$fit(multistart=10, trace=TRUE)

# Instantiate and save log-likelihood emulator object.
llik_em <- llikEmulatorGP("llik_em", gp_obj, default_conditional=FALSE, 
                          default_normalize=TRUE, lik_par=inv_prob$sig2_model, 
                          llik_bounds=llik_bounds)
```

### Run approximate MCMC
There are various GP-approximated MCMC algorithms implemented in 
`gp_mcmc_functions.r`. We keep things simple here and simply use the GP 
plug-in mean approach; i.e., replace the true log-likelihood with the GP 
predictive mean.

```{r}
mcmc_settings_mean <- list(test_label="mean_round1", approx_type="mean",
                           mcmc_func_name="mcmc_bt_wrapper", 
                           sampler="DEzs", n_itr=50000L, try_parallel=FALSE,  
                           n_chain=4L, defer_ic=TRUE)

# MCMC sampling using exact likelihood: with and without prior truncation.
mcmc_list_mean <- run_mcmc(llik_em, par_prior, mcmc_settings_mean)

# Store samples.
burn_in_mean <- round(mcmc_settings_mean$n_itr/2)
samp_dt_mean <- select_mcmc_itr(mcmc_list_mean$samp, itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_mean)
info_dt <- select_mcmc_itr(mcmc_list_mean$info, itr_start=burn_in_mean)
```

```{r}
trace_plts_mean <- get_trace_plots(samp_dt, test_label="mean")
for(plt in trace_plts_mean) plot(plt)
```

```{r}
hist_plts_mean <- get_hist_plot_comparisons(samp_dt, test_label_baseline="exact")
for(plt in hist_plts_mean) plot(plt)
```

## EKI
```{r}
# Run single step of EKI.
Sig <- diag(llik_em$lik_par, nrow=inv_prob$dim_obs)
eki_output <- run_eki(inv_prob$y, inv_prob$par_to_obs_op, 
                      Sig, par_prior=par_prior, U0=design_info$input, 
                      G0=design_info$fwd, n_itr=1)

samp_dt <- append_samples_mat(samp_dt, eki_output$U, param_type="par",
                              test_label="eki_1_itr")
```


```{r}
hist_plts_eki <- get_hist_plot_comparisons(samp_dt, test_label_baseline="exact", 
                                           test_labels="eki_1_itr")
for(plt in hist_plts_eki) plot(plt)
```

## Compare the two approximations
```{r}
kde_plts <- get_1d_kde_plots(samp_dt, test_label_baseline="exact")
for(plt in kde_plts) plot(plt)
```

# Fitting GP Emulator using EKI Ensemble as Design
We now consider fitting a GP emulator, where the approximate EKI samples are 
used as the initial design for the GP. The idea is that EKI "zooms in" on the 
important region of parameter space, which allows for a GP fit that is more 
fine-tuned to approximate the log-likelihood well in the region of significant
posterior mass. The cost of this is another round of forward model runs, as the 
forward model needs to be run again at the EKI samples to generate the 
outputs for fitting the GP. TODO: should really be combining the EKI initial 
and posterior ensemble as the design for the GP.

## Fit GP
```{r}
# Execute forward model runs at EKI points.
design_info_eki <- get_init_design_list(inv_prob, N_design=n_design, 
                                        design_method="eki", inputs=eki_output$U)

# Fit GP for log-likelihood.
gp_obj_eki <- gpWrapperKerGP(design_info_eki$input, matrix(design_info_eki$llik, ncol=1), 
                             scale_input=TRUE, normalize_output=TRUE)
gp_obj_eki$set_gp_prior("Gaussian", "quadratic", include_noise=FALSE)
gp_obj_eki$fit(multistart=10, trace=TRUE)

# Instantiate and save log-likelihood emulator object.
llik_em_eki <- llikEmulatorGP("llik_em_eki", gp_obj_eki, default_conditional=FALSE, 
                              default_normalize=TRUE, lik_par=inv_prob$sig2_model, 
                              llik_bounds=llik_bounds)
```

## Compare the Two GP Fits

### Evaluate on Test Points Sampled from Prior
```{r}
em_pred_list <- llik_em$predict_emulator(test_info_prior$input, return_var=TRUE)
em_pred_list_eki <- llik_em_eki$predict_emulator(test_info_prior$input, return_var=TRUE)
```

```{r}
llik_em$plot_pred_validation(test_info_prior$input, test_info_prior$llik,
                             em_pred_list=em_pred_list, 
                             plot_title="llik_em predictions: prior test points")

llik_em_eki$plot_pred_validation(test_info_prior$input, test_info_prior$llik,
                                 em_pred_list=em_pred_list_eki,
                                 plot_title="llik_em_eki predictions: prior test points")
```

### Evaluate on Test Points Sampled from True Posterior
```{r}
em_pred_list_post <- llik_em$predict_emulator(test_info_exact$input, return_var=TRUE)
em_pred_list_eki_post <- llik_em_eki$predict_emulator(test_info_exact$input, return_var=TRUE)
```

```{r}
# Plotting Gaussian predictions, along with llik bound. 

llik_em$plot_pred_validation(test_info_exact$input, test_info_exact$llik,
                             em_pred_list=em_pred_list_post, adjustment="none",
                             plot_title="llik_em predictions: true posterior test points") +
                             geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")

llik_em_eki$plot_pred_validation(test_info_exact$input, test_info_exact$llik,
                                 em_pred_list=em_pred_list_eki_post, adjustment="none",
                                 plot_title="llik_em_eki predictions: true posterior test points") + 
                                 geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")
```
```{r}
# Plotting truncated Gaussian predictions, upper truncated using llik bound. 

llik_em$plot_pred_validation(test_info_exact$input, test_info_exact$llik,
                             em_pred_list=em_pred_list_post, adjustment="truncated",
                             plot_title="llik_em predictions: true posterior test points") +
                             geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")

llik_em_eki$plot_pred_validation(test_info_exact$input, test_info_exact$llik,
                                 em_pred_list=em_pred_list_eki_post, adjustment="truncated",
                                 plot_title="llik_em_eki predictions: true posterior test points") + 
                                 geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")
```

## Run approximate posterior sampling using new GP emulator.

```{r}
mcmc_settings_mean_eki <- list(test_label="mean_eki", approx_type="mean",
                           mcmc_func_name="mcmc_bt_wrapper", 
                           sampler="DEzs", n_itr=50000L, try_parallel=FALSE,  
                           n_chain=4L, defer_ic=TRUE)

# MCMC sampling using exact likelihood: with and without prior truncation.
mcmc_list_mean_eki <- run_mcmc(llik_em_eki, par_prior, mcmc_settings_mean_eki)

# Store samples.
samp_dt_mean_eki <- select_mcmc_itr(mcmc_list_mean_eki$samp, 
                                    itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_mean_eki)
```

```{r}
trace_plts_mean_eki <- get_trace_plots(samp_dt, test_label="mean_eki")
for(plt in trace_plts_mean_eki) plot(plt)
```

```{r}
hist_plts_mean_eki <- get_hist_plot_comparisons(samp_dt, test_labels="mean_eki",
                                                test_label_baseline="exact")
for(plt in hist_plts_mean_eki) plot(plt)
```

## Compare theposterior approximations
```{r}
kde_plts <- get_1d_kde_plots(samp_dt, test_label_baseline="exact")
for(plt in kde_plts) plot(plt)
```

# Multi-Step EKI 
We now consider running multiple steps of EKI to generate a design for the 
GP. Splitting the EKI update into multiple steps via likelihood tempering tends
to produce a much better approximation than single-step EKI.

```{r}
# Run multi-step EKI.
n_itr_eki <- 3L
eki_multistep_output <- run_eki(inv_prob$y, inv_prob$par_to_obs_op, 
                                Sig, par_prior=par_prior, U0=design_info$input, 
                                G0=design_info$fwd, n_itr=n_itr_eki)

samp_dt <- append_samples_mat(samp_dt, eki_multistep_output$U, param_type="par",
                              test_label=paste0("eki_", n_itr_eki, "_itr"))
```

```{r}
kde_plts <- get_1d_kde_plots(samp_dt, test_label_baseline="exact")
for(plt in kde_plts) plot(plt)
```
## Fit GP on samples from Multi-step EKI
```{r}
# Execute forward model runs at EKI points.
design_info_eki_multi <- get_init_design_list(inv_prob, N_design=n_design, 
                                              design_method=paste0("eki_", n_itr_eki), 
                                              inputs=eki_multistep_output$U)

# Fit GP for log-likelihood.
gp_obj_eki_multi <- gpWrapperKerGP(design_info_eki_multi$input, 
                                   matrix(design_info_eki_multi$llik, ncol=1), 
                                   scale_input=TRUE, normalize_output=TRUE)
gp_obj_eki_multi$set_gp_prior("Gaussian", "quadratic", include_noise=FALSE)
gp_obj_eki_multi$fit(multistart=10, trace=TRUE)

# Instantiate and save log-likelihood emulator object.
llik_em_eki_multi <- llikEmulatorGP("llik_em_eki_multi", gp_obj_eki_multi, 
                                    default_conditional=FALSE, 
                                    default_normalize=TRUE, 
                                    lik_par=inv_prob$sig2_model, 
                                    llik_bounds=llik_bounds)
```


## Approximate Posterior Sampling
```{r}
test_label_eki_multi <- paste0("mean_eki_", n_itr_eki)
mcmc_settings_mean_eki_multi <- list(test_label=test_label_eki_multi, 
                                     approx_type="mean",
                                     mcmc_func_name="mcmc_bt_wrapper", 
                                     sampler="DEzs", n_itr=50000L, 
                                     try_parallel=FALSE,  
                                     n_chain=4L, defer_ic=TRUE)

# MCMC sampling using exact likelihood: with and without prior truncation.
mcmc_list_mean_eki_multi <- run_mcmc(llik_em_eki_multi, par_prior, 
                                     mcmc_settings_mean_eki_multi)

# Store samples.
samp_dt_mean_eki_multi <- select_mcmc_itr(mcmc_list_mean_eki_multi$samp, 
                                          itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_mean_eki_multi)
```


```{r}
trace_plts_mean_eki_multi <- get_trace_plots(samp_dt, test_label=test_label_eki_multi)
for(plt in trace_plts_mean_eki_multi) plot(plt)
```


```{r}
hist_plts_mean_eki_multi <- get_hist_plot_comparisons(samp_dt, 
                                                      test_labels=test_label_eki_multi,
                                                      test_label_baseline="exact")
for(plt in hist_plts_mean_eki_multi) plot(plt)
```

```{r}
kde_plts <- get_1d_kde_plots(samp_dt, test_label_baseline="exact")
for(plt in kde_plts) plot(plt)
```

Thoughts on what happened here: the 3-step EKI posterior samples were quite 
concentrated in the parameter space, and reflected the true posterior quite 
well. So the GP emulator probably predicts the posterior density quite well 
in the region of significant posterior mass, but is quite uncertain elsewhere.
In this case I'm assuming it is way overpredicting the posterior density in 
regions where it is insignificant, leading to a very flat approximate posterior.
Should confirm this is what's happening. It would seem that training the GP 
on all of the EKI samples (not just the final round) would help a lot here, as 
this would include samples from regions of insignificant posterior mass.














