---
title: "llikEmulator Class"
author: "Andrew Roberts"
date: "2024-11-27"
output: html_document
---

This file provides an introduction to the `llikEmulator` class and associated 
classes that inherit from this base class. The code implementing these 
classes can be found in `llikEmulator.r`.

# Introduction
The `llikEmulator` class encapsulates a surrogate (i.e., emulator) model for 
a log-likelihood function. It is designed to support both stochastic 
and deterministic log-likelihood surrogates. Users do not typically work 
directly with the base `llikEmulator` class, but rather with objects from 
classes that inherit from `llikEmulator`. For example, the `llikEmulatorGP`
class is used in the case that a log-likelihood is emulated by a Gaussian 
process (GP). In this example, `llikEmulatorGP` is agnostic to the specific 
distributional form of the underlying log-likelihood. Other classes are designed
for specific likelihood structures; for example, `llikEmulatorGPFwdGauss` 
handles the specific case of a Gaussian likelihood 
$\mathcal{N}(\mathcal{G}(u), \Sigma)$ where the forward model 
$\mathcal{G}(\cdot)$ is replaced by a GP emulator. Having more specific classes 
to handle common special cases allows one to leverage the additional structure 
in certain contexts.

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo=TRUE)

library(data.table)

base_dir <- file.path("/projectnb", "dietzelab", "arober", "gp-calibration")
src_dir <- file.path(base_dir, "src")

# Source required files.
source(file.path(src_dir, "gpWrapper.r"))
source(file.path(src_dir, "llikEmulator.r"))
source(file.path(src_dir, "general_helper_functions.r"))
source(file.path(src_dir, "inv_prob_test_functions.r"))
source(file.path(src_dir, "seq_design.r"))
source(file.path(src_dir, "statistical_helper_functions.r"))
source(file.path(src_dir, "plotting_helper_functions.r"))
source(file.path(src_dir, "gp_helper_functions.r"))
```

# Basic Functionality 
We start by instantiating an object of class `llikEmulatorGP`, meaning that 
we will consider directly emulating the log-likelihood with a GP. This class 
is designed to work with GP objects from `gpWrapper` classes. We start by 
setting up a pre-defined example inverse problem, fit a GP approximating the 
log-likelihood map, and then define a `llikEmulatorGP` object. 

```{r}
# Example inverse problem, with associated log-likelihood. This inverse problem 
# is based on calibrating parameters of the VSEM vegetation model.
inv_prob <- get_vsem_test_1()
```

```{r}
# Create a design: input-output log-likelihood pairs, the training data for 
# the GP model.
design_info <- get_init_design_list(inv_prob, design_method="LHS", N_design=250L)

# Assemble test/validation points that will be used later on.
test_info <- get_init_design_list(inv_prob, design_method="LHS", N_design=500L)
```

```{r}
# Train a GP model on the design points.
gp_obj <- gpWrapperKerGP(design_info$input, matrix(design_info$llik, ncol=1), 
                         normalize_output=TRUE, scale_input=TRUE)
gp_obj$fit("Gaussian_plus_Quadratic", "constant", estimate_nugget=FALSE, 
           optimFun="nloptr::nloptr", trace=TRUE, multistart=10)
```
```{r}
# Create log-likelihood surrogate object.
llik_em <- llikEmulatorGP("em_llik", gp_obj, 
                          lik_par=inv_prob$sig2_model, use_fixed_lik_par=TRUE,
                          default_conditional=FALSE, default_normalize=TRUE)
```

## Log-likelihood and Likelihood Predictive Distribution

## Deterministic Functions of the Likelihood Surrogate
The `llikEmulator` class offers a standardized interface for computing 
deterministic functions of the (potentially stochastic) likelihood surrogate.
The major use case is computing deterministic approximations of the likelihood; 
this is handled by the method `calc_lik_approx()`.  
Note that some likelihood approximations will only be 
available for certain specific `llikEmulator` classes, while others are 
general enough that they can be implemented generally in the base `llikEmulator`
class.

These functions are designed to operate on the `em_pred_list` object returned 
by the method `predict_emulator()`; i.e., the perspective here is to view 
quantities of interest to be some function of the underlying emulator 
predictive distribution. The methods shown below will automatically compute 
`em_pred_list` under the hood, but if the predictions at a set of inputs are 
going to be used for a variety of purposes, it is recommended to pre-compute 
`em_pred_list` and pass them to subsequent calls to avoid re-computing 
predictions each time.

```{r}
# Emulator predictions.
em_pred_list <- llik_em$predict_emulator(input=test_info$input)
```

### Deterministic likelihood approximations
`calc_lik_approx()` provides a high-level convenience function for computing 
deterministic approximations of the likelihood function on an input-by-input 
basis; i.e., pointwise approximation of the likelihood. The first argument 
`approx_type` specifies the specific approximation to use. The method will 
look for other methods with the name `calc_<approx_type>_approx()`. By default, 
the log of the approximations are returned for numerical stability. The 
approximations on the original scale can be returned by setting the argument 
`log_scale = TRUE`.

```{r}
# Compute plug-in mean approximation.
approx_mean <- llik_em$calc_lik_approx("mean", em_pred_list=em_pred_list,
                                       simplify=TRUE)

# Compute marginal approximation (expectation of likelihood wrt emulator).
approx_marg <- llik_em$calc_lik_approx("marginal", em_pred_list=em_pred_list,
                                       simplify=TRUE)

head(cbind(mean=approx_mean, marginal=approx_marg))
```
The method `calc_lik_approx()` actually allows the first argument `approx_type`
to be a vector, so that the above two approximations can be computed in a 
single call.

```{r}
# Compute both plug-in mean and marginal approximation.
approx_mean_marg <- llik_em$calc_lik_approx(c("mean", "marginal"), 
                                            em_pred_list=em_pred_list,
                                            return_type="matrix")

head(approx_mean_marg)
```
The deterministic approximate log-likelihood can also be returned as a function.
This is useful for interfacing with external code; for example, MCMC 
implementations that require a log-likelihood function as input.

```{r}
# Store approximate log-likelihood function.
llik_approx <- llik_em$get_llik_func("mean")

# Evaluate log-likelihood function at a set of inputs.
llik_approx_vals <- llik_approx(test_info$input)
head(llik_approx_vals)
```

### Error Metrics
The method `calc_lik_approx_pw_err()` provides a convenient high-level
interface for computing deterministic likelihood predictions and evaluating 
their error with respect to some baseline true log-likelihood values. Here, 
"pw" means "pointwise", as the errors are computed point-by-point. The method 
allows computing errors for multiple likelihood approximations and multiple 
error metrics in a single call. 

```{r}
# Compute pointwise squared errors, absolute errors, and their "weighted" 
# versions, which are divided by the log-likelihood emulator predictive  
# standard deviation.
errs_mean_marg <- llik_em$calc_lik_approx_pw_err(llik_true=test_info$llik, 
                                                 err_type=c("mse", "wmse", "mae", "wmae"), 
                                                 llik_pred=approx_mean_marg, 
                                                 em_pred_list=em_pred_list,
                                                 return_type="data.table")

head(errs_mean_marg)
```


### Generic Functions
Both `calc_lik_approx()` and `calc_err_metric()` are wrappers around the 
lower-level method `calc_func()`, which allows a user to supply a function 
that operates on `em_pred_list`. We provide an example of this below.
```{r}
# User-defined function.
f <- function(em_pred_list, ...) em_pred_list$mean + 2*sqrt(em_pred_list$var)

# Apply function to `em_pred_list`.
f_test <- llik_em$calc_func(func=f, em_pred_list=em_pred_list)
head(f_test)
```

The method `calc_multi_func()` is a wrapper around `calc_func()` that allows 
for applying multiple functions to the same `em_pred_list` object. This is 
actually the lower-level method directly called by `calc_lik_approx()` and 
`calc_err_metric()`, which allows multiple approximation methods/error metrics
to be passed. By default, `calc_multi_func()` returns a list, but setting 
`return_type = "matrix"` causes a matrix to be returned, with one column per 
function in the list.

```{r}
# List of two functions to evaluate.
g <- function(em_pred_list, ...) em_pred_list$mean + 4*sqrt(em_pred_list$var)
func_list <- list(f=f, g=g)

# Apply each function to `em_pred_list`.
func_tests <- llik_em$calc_multi_func(func_list=func_list, 
                                      em_pred_list=em_pred_list, 
                                      return_type="matrix")
head(func_tests)
```
