---
title: "Tutorial: Helper Functions to Manipulate and Plot MCMC Output"
author: "Andrew Roberts"
date: "2024-10-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)

library(ggplot2)
library(data.table)
library(assertthat)

base_dir <- file.path("/projectnb", "dietzelab", "arober", "gp-calibration")
src_dir <- file.path(base_dir, "src")

source(file.path(src_dir, "mcmc_helper_functions.r"))
source(file.path(src_dir, "general_helper_functions.r"))
```

# Creating `samp_dt` object.
Let's consider an example where we are comparing 2 sampling algorithms, which 
we will label "test1" and "test2", respectively. The first has two chains and 
two parameter types (groups), which we will call "group1" and "group2". The 
first group will have two parameters called "par1" and "par2" while the second 
group has a single parameter called "par". "test2" will be the same except 
only with one chain and the first parameter type.

## Creating some samples for example.
Note that samples will typically come from an MCMC algorithm, but we are keeping
things simple here and sampling from Gaussians.
```{r}
set.seed(236643)

n_samp <- 1000
m1 <- c(0,0)
C1 <- matrix(c(1,.8,.8, 1), nrow=2)

# Helper functions for simulating some samples.
samp_group1 <- function(m, C, n) {
  Z <- matrix(rnorm(2*n, 0, 1), nrow=2, ncol=n)
  L <- t(chol(C))
  samp_mat <- add_vec_to_mat_rows(m, t(L %*% Z))
  colnames(samp_mat) <- c("par1","par2")
  return(samp_mat)
}

samp_group2 <- function(n) {
  samp_mat <- matrix(rnorm(n), ncol=1)
  colnames(samp_mat) <- "par"
  return(samp_mat)
}

# Generate samples for test1.
test1_chain1 <- list(type1 = samp_group1(m1, C1, n_samp),
                     type2 = samp_group2(n_samp))
test1_chain2 <- list(type1 = samp_group1(m1, C1, n_samp),
                     type2 = samp_group2(n_samp))
test1_chain_list <- list(test1_chain1, test1_chain2)

# Generate samples for test2.
shift <- c(.5,.5)
test2_list <- list(type1 = samp_group1(m1+shift, C1, n_samp))
```

## Generating the `samp_dt` object.
```{r}
# Start by creating the samples data.table object for test1.
samp_dt <- format_mcmc_output_multi_chain(test1_chain_list, test_label="test1")

# Now we append the samples from test2, which only has one chain. If this 
# list used multiple chains we would have to use 
# `append_mcmc_output_multi_chain()` instead.
samp_dt <- append_mcmc_output(samp_dt, test2_list, test_label="test2")

# Can use this function to check that an R object satisfies the requirements 
# to be a data.table samples object.
assert_is_samp_dt(samp_dt)

# Printing number of samples in each category.
by_cols <- c("test_label", "param_type", "param_name", "chain_idx")
samp_dt[,.N, by=by_cols]
```

# Subsetting the `samp_dt` data.table.
The main function for subsetting is `select_mcmc_samp()`. It subsets the rows 
of `samp_dt` by selecting specified values of test labels, parameter types, 
parameter names, chain indices, and iteration ranges. Iteration ranges can be 
selected by test label.
```{r}
# Example 1: select iterations in [200,500] from all parameters in all test 
# labels from only chains with index 1 and in param type 1.
samp_dt_subset1 <- select_mcmc_samp(samp_dt, itr_start=200L, itr_stop=500L,
                                    param_types="type1", chain_idcs=1L)
print(samp_dt_subset1[,.(min_itr=min(itr), max_itr=max(itr)), by=by_cols])

# Example 2: Define different iteration range for different test labels.
itr_start <- c(test1=100L, test2=900L)
itr_stop <- c(test1=200L, test2=1000L)
samp_dt_subset2 <- select_mcmc_samp(samp_dt, itr_start=itr_start, itr_stop=itr_stop)
print(samp_dt_subset2[,.(min_itr=min(itr), max_itr=max(itr)), by=by_cols])
```


There is also a similar function that returns the samples for a specific 
test label/parameter type combination in a typical matrix format. Samples 
belonging to different chains may each be returned in their matrix, or all 
chains can be combined into a single matrix. The `param_names` argument 
here is useful for controlling the order of the columns in the returned matrices.

```{r}
# Example 1: return list of matrices, one per chain. Explicitly passing the 
# parameter names to emphasize the effect on the column order of the matrices.
# Rownames are set to iteration names.
samp_mat_list <- select_mcmc_samp_mat(samp_dt, test_label="test1", param_type="type1",
                                      param_names=c("par2","par1"), itr_start=1L, 
                                      itr_stop=6L, return_chain_list=TRUE)
print(samp_mat_list$`1`) # Chain 1
print(samp_mat_list$`2`) # Chain 2

# Example 2: combine into single matrix. Rownames are of the form 
# "<chain_idx>_<itr>".
samp_mat <- select_mcmc_samp_mat(samp_dt, test_label="test1", param_type="type1",
                                 param_names=c("par2","par1"), itr_start=1L, 
                                 itr_stop=6L, return_chain_list=FALSE)
print(samp_mat)
```

# Plotting
The plot functions all use the same row selection/subsetting interface as 
`select_mcmc_samp()`, which is used to control which test labels, parameters, etc.
are plotted. 

## Trace Plots
A single parameter is plotted on each plot. Different chains can be overlaid 
on the same plot, or plotted on separate plots.
```{r}
# Example 1: overlay chains on single plot.
trace1 <- get_trace_plots(samp_dt, test_labels="test1", param_type="type1",
                          itr_start=500L, overlay_chains=TRUE)
for(plt in trace1) plot(plt)

# Example 2: each chain gets a separate plot.
trace2 <- get_trace_plots(samp_dt, test_labels="test1", param_type="type1",
                          itr_start=500L, overlay_chains=FALSE)
for(plt in trace2) plot(plt)
```

## Histograms
```{r}
# Example 1: Combining samples across all chains.
hist1 <- get_hist_plots(samp_dt, param_types="type1", combine_chains=TRUE, bins=30)
for(plt in hist1) plot(plt)

# Example 2: Same thing but with frequency polygon instead of histogram.
hist2 <- get_hist_plots(samp_dt, param_types="type1", combine_chains=TRUE, 
                        plot_type="freqpoly", bins=30)
for(plt in hist2) plot(plt)

# Example 3: Keeping chains separate. Only looking at first test label here to 
# avoid having too many histograms on the same plot.
hist3 <- get_hist_plots(samp_dt, test_labels="test1", param_types="type1", 
                        chain_idcs=NULL, combine_chains=FALSE, bins=30)
for(plt in hist3) plot(plt)
```






