---
title: "Comparing Approximations of Likelihood Ratio"
author: "Andrew Roberts"
date: "2025-01-23"
output: html_document
---

This file assesses the ability to emulate the log-likelihood surface in an
inverse problem using a Gaussian process (GP). In particular, it investigates
the challenge of emulating a log-likelihood that exhibits a large dynamic 
range, which can prove troublesome for stationary GP priors. GP predictive 
accuracy is assessed with respect to the prior distribution, as well as with 
respect to the true posterior distribution of the Bayesian inverse problem.
Often, the posterior is highly concentrated relative to the prior, and the 
GP fit might look reasonable on a global scale, but be completely off in the 
region of true posterior support.

- maybe this problem is just quite hard.
- maybe its okay the mean prediction looks bad after the first round; that is 
what sequential design is for. What's more concerning is the MCMC runs not even
working; it seems that the truncated approach seems to reduce these errors.
- this is a situation where sub-sampling the current approx posterior might not
perform well, but more targeted acquisition functions should hopefully help.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)

library(data.table)
library(ggplot2)

# Filepath definitions.
base_dir <- file.path("/projectnb", "dietzelab", "arober", "gp-calibration")
src_dir <- file.path(base_dir, "src")

# Source required files.
source(file.path(src_dir, "general_helper_functions.r"))
source(file.path(src_dir, "inv_prob_test_functions.r"))
source(file.path(src_dir, "statistical_helper_functions.r"))
source(file.path(src_dir, "plotting_helper_functions.r"))
source(file.path(src_dir, "seq_design.r"))
source(file.path(src_dir, "gp_helper_functions.r"))
source(file.path(src_dir, "gpWrapper.r"))
source(file.path(src_dir, "llikEmulator.r"))
source(file.path(src_dir, "mcmc_helper_functions.r"))
source(file.path(src_dir, "gp_mcmc_functions.r"))
```

# Setup: load inverse problem
```{r}
# Load inverse problem setup information. Store log-likelihood upper bound.
inv_prob <- get_vsem_test_1()
par_names <- inv_prob$par_names
par_prior <- inv_prob$par_prior
llik_bounds <- inv_prob$llik_obj$get_llik_bounds()

print(paste0("Input dimension: ", inv_prob$dim_par))
print(paste0("Observation space dimension: ", inv_prob$dim_obs))
print(paste0("Log-likelihood bounds: ", paste(llik_bounds, collapse=", ")))
```

```{r}
# Create initial design.
n_design <- 250L
design_method <- "LHS"
design_info <- get_init_design_list(inv_prob, design_method, n_design)

print(paste0("Number design points: ", n_design))
print(paste0("Design method: ", design_method))
```

```{r}
# Sample test points from prior.
n_design_test <- 500L
test_info_prior <- get_init_design_list(inv_prob, design_method, n_design_test)
```


## 1d projections of llik surface

### Fixed values at midpoints of design bounds.
```{r}
# 1d projections of llik surface.
inv_prob$llik_obj$plot_1d_projection(input_bounds=design_info$bounds)
```
### Fixed values randomly sampled.
```{r}
n_random_samp <- 5
fixed_inputs <- sample_prior(par_prior, n_random_samp)

for(i in seq_len(n_random_samp)) {
  plts <- inv_prob$llik_obj$plot_1d_projection(input_fixed=fixed_inputs[i,,drop=FALSE], 
                                               input_bounds=design_info$bounds)
  for(plt in plts) plot(plt)
}
```


# Exact Posterior Samples
```{r}
# Run exact MCMC.
mcmc_settings <- list(test_label="exact", mcmc_func_name="mcmc_bt_wrapper", 
                      sampler="DEzs", n_itr=50000L, try_parallel=FALSE,  
                      n_chain=4L, defer_ic=TRUE)

# MCMC sampling using exact likelihood.
mcmc_list <- run_mcmc(inv_prob$llik_obj, par_prior, mcmc_settings)

# Table storing MCMC samples.
burn_in_exact <- round(mcmc_settings$n_itr/2)
samp_dt <- select_mcmc_itr(mcmc_list$samp, itr_start=burn_in_exact)
mcmc_metadata <- mcmc_list$output_list
```

```{r}
trace_plts <- get_trace_plots(samp_dt)
for(plt in trace_plts) plot(plt)
```
```{r}
# 1d marginals, compared to prior.
prior_samp <- sample_prior(par_prior, n=1e6)
samp_dt_temp <- append_samples_mat(samp_dt, prior_samp, param_type="par", 
                                   test_label="prior")

hist_plts_exact <- get_hist_plot_comparisons(samp_dt_temp, test_label_baseline="prior")
for(plt in hist_plts_exact) plot(plt)

rm(list="samp_dt_temp")
```

```{r}
# 2d marginals, compared to prior.
kde2d_plts_exact <- get_2d_density_plots(samp_dt)

for(plt in kde2d_plts_exact$exact) plot(plt)
```


```{r}
# Test points subsampled from true posterior.
samp_mat <- select_mcmc_samp_mat(samp_dt, test_label="exact", 
                                 param_type="par")[,par_names]
test_info_exact <- get_init_design_list(inv_prob, design_method="subsample", 
                                        N_design=n_design_test,
                                        design_candidates=samp_mat)
```


# Assessing Quadratic Fit to Log-Likelihood Surface

## Quadratic Fit, no cross terms
```{r}
df <- data.frame(y=design_info$llik, design_info$input)
quad_formula <- as.formula(paste0("y ~ ", paste0("poly(", inv_prob$par_names, ", degree=2)", 
                                                 collapse=" + ")))

lm_quad <- lm(quad_formula, df)
resid_quad <- resid(lm_quad)
pred_quad_design <- predict(lm_quad)
pred_quad_prior <- predict(lm_quad, newdata=data.frame(test_info_prior$input))
pred_quad_exact <- predict(lm_quad, newdata=data.frame(test_info_exact$input))

coef(lm_quad)
```

```{r}
plot(design_info$llik, pred_quad_design, main="Predictions at design", 
     xlab="true", ylab="predicted")
abline(a=0,b=1, col="red")
abline(h=llik_bounds[2], col="blue", lty="dashed")
```

```{r}
plot(test_info_prior$llik, pred_quad_prior, main="Predictions at prior points", 
     xlab="true", ylab="predicted")
abline(a=0,b=1, col="red")
abline(h=llik_bounds[2], col="blue", lty="dashed")
```

```{r}
plot(test_info_exact$llik, pred_quad_exact, main="Predictions at posterior points", 
     xlab="true", ylab="predicted")
abline(a=0,b=1, col="red")
abline(h=llik_bounds[2], col="blue", lty="dashed")
```

## Quadratic Fit, cross terms

```{r}
B <- poly(design_info$input, degree=2)
df2 <- data.frame(y=design_info$llik, B)
lm_quad2 <- lm(y~., df2)
resid_quad2 <- resid(lm_quad2)
pred_quad2_design <- predict(lm_quad2)

B_prior <- poly(test_info_prior$input, degree=2)
pred_quad2_prior <- predict(lm_quad2, newdata=data.frame(B_prior))

B_exact <- poly(test_info_exact$input, degree=2)
pred_quad2_exact <- predict(lm_quad2, newdata=data.frame(B_exact))

coef(lm_quad2)
```

```{r}
plot(design_info$llik, pred_quad2_design, main="Predictions at design", 
     xlab="true", ylab="predicted")
abline(a=0,b=1, col="red")
abline(h=llik_bounds[2], col="blue", lty="dashed")
```

```{r}
plot(test_info_prior$llik, pred_quad2_prior, main="Predictions at prior points", 
     xlab="true", ylab="predicted")
abline(a=0,b=1, col="red")
abline(h=llik_bounds[2], col="blue", lty="dashed")
```

```{r}
plot(test_info_exact$llik, pred_quad2_exact, main="Predictions at posterior points", 
     xlab="true", ylab="predicted")
abline(a=0,b=1, col="red")
abline(h=llik_bounds[2], col="blue", lty="dashed")
```

# Fit Log-Likelihood GP Emulators

## Gaussian Kernel, Quadratic Mean (no cross terms)
```{r}
# Fit GP for log-likelihood.
gp_obj <- gpWrapperKerGP(design_info$input, matrix(design_info$llik, ncol=1), 
                         scale_input=TRUE, normalize_output=TRUE)
gp_obj$set_gp_prior("Gaussian", "quadratic", include_noise=FALSE)
gp_obj$fit(multistart=10, trace=TRUE)

# Instantiate and save log-likelihood emulator object.
llik_em <- llikEmulatorGP("llik_em", gp_obj, default_conditional=FALSE, 
                          default_normalize=TRUE, lik_par=inv_prob$sig2_model, 
                          llik_bounds=llik_bounds)
```


```{r}
# Compute emulator predictions.
em_pred_list_prior <- llik_em$predict_emulator(test_info_prior$input)
em_pred_list_exact <- llik_em$predict_emulator(test_info_exact$input)
```

### Gaussian Predictions
```{r}
interval_method <- "pm_std_dev"
N_std_dev <- 2

llik_plot_prior <- llik_em$plot_pred_validation(input=test_info_prior$input,
                                                true_llik=test_info_prior$llik,
                                                em_pred_list=em_pred_list_prior, 
                                                interval_method=interval_method,
                                                N_std_dev=N_std_dev, adjustment="none", 
                                                plot_title="Prior Predictions") + 
                   geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")

llik_plot_exact<- llik_em$plot_pred_validation(input=test_info_exact$input,
                                                true_llik=test_info_exact$llik,
                                                em_pred_list=em_pred_list_exact, 
                                                interval_method=interval_method,
                                                N_std_dev=N_std_dev, adjustment="none", 
                                                plot_title="Posterior Predictions") + 
                   geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")

plot(llik_plot_prior)
plot(llik_plot_exact)
```

### Truncated Gaussian Predictions
```{r}
llik_plot_prior <- llik_em$plot_pred_validation(input=test_info_prior$input,
                                                true_llik=test_info_prior$llik,
                                                em_pred_list=em_pred_list_prior, 
                                                interval_method=interval_method,
                                                N_std_dev=N_std_dev, adjustment="truncated", 
                                                plot_title="Prior Predictions") + 
                   geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")

llik_plot_exact <- llik_em$plot_pred_validation(input=test_info_exact$input,
                                                true_llik=test_info_exact$llik,
                                                em_pred_list=em_pred_list_exact, 
                                                interval_method=interval_method,
                                                N_std_dev=N_std_dev, adjustment="truncated", 
                                                plot_title="Posterior Predictions") + 
                   geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")

plot(llik_plot_prior)
plot(llik_plot_exact)
```

## Gaussian Plus Quadratic Kernel, Constant Mean
```{r}
# Fit GP for log-likelihood.
gp_obj2 <- gpWrapperKerGP(design_info$input, matrix(design_info$llik, ncol=1), 
                          scale_input=TRUE, normalize_output=TRUE)
gp_obj2$set_gp_prior("Gaussian_plus_Quadratic", "constant", include_noise=FALSE)
gp_obj2$fit(multistart=10, trace=TRUE)

# Instantiate and save log-likelihood emulator object.
llik_em2 <- llikEmulatorGP("llik_em2", gp_obj2, default_conditional=FALSE, 
                          default_normalize=TRUE, lik_par=inv_prob$sig2_model, 
                          llik_bounds=llik_bounds)
```

```{r}
# Compute emulator predictions.
em2_pred_list_prior <- llik_em2$predict_emulator(test_info_prior$input)
em2_pred_list_exact <- llik_em2$predict_emulator(test_info_exact$input)
```

### Gaussian Predictions
```{r}
llik_plot_prior <- llik_em2$plot_pred_validation(input=test_info_prior$input,
                                                true_llik=test_info_prior$llik,
                                                em_pred_list=em2_pred_list_prior, 
                                                interval_method=interval_method,
                                                N_std_dev=N_std_dev, adjustment="none", 
                                                plot_title="Prior Predictions") + 
                   geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")

llik_plot_exact <- llik_em2$plot_pred_validation(input=test_info_exact$input,
                                                true_llik=test_info_exact$llik,
                                                em_pred_list=em2_pred_list_exact, 
                                                interval_method=interval_method,
                                                N_std_dev=N_std_dev, adjustment="none", 
                                                plot_title="Posterior Predictions") + 
                   geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")

plot(llik_plot_prior)
plot(llik_plot_exact)
```

### Truncated Gaussian Predictions
```{r}
llik_plot_prior <- llik_em2$plot_pred_validation(input=test_info_prior$input,
                                                true_llik=test_info_prior$llik,
                                                em2_pred_list=em_pred_list_prior, 
                                                interval_method=interval_method,
                                                N_std_dev=N_std_dev, adjustment="truncated", 
                                                plot_title="Prior Predictions") + 
                   geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")

llik_plot_exact <- llik_em2$plot_pred_validation(input=test_info_exact$input,
                                                true_llik=test_info_exact$llik,
                                                em2_pred_list=em_pred_list_exact, 
                                                interval_method=interval_method,
                                                N_std_dev=N_std_dev, adjustment="truncated", 
                                                plot_title="Posterior Predictions") + 
                   geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")

plot(llik_plot_prior)
plot(llik_plot_exact)
```

# Plug-In Mean Approximation: MCMC Sampling

## Using Truncated Predictive Distribution
```{r}
mcmc_settings_mean <- list(test_label="mean", approx_type="mean",
                           adjustment="truncated",
                           mcmc_func_name="mcmc_bt_wrapper", 
                           sampler="DEzs", n_itr=50000L, try_parallel=FALSE,  
                           n_chain=4L, defer_ic=TRUE)

# MCMC sampling using exact likelihood: with and without prior truncation.
mcmc_list_mean <- run_mcmc(llik_em, par_prior, mcmc_settings_mean)

# Store samples.
burn_in_mean <- round(mcmc_settings_mean$n_itr/2)
samp_dt_mean <- select_mcmc_itr(mcmc_list_mean$samp, itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_mean)
info_dt <- select_mcmc_itr(mcmc_list_mean$info, itr_start=burn_in_mean)
```

```{r}
trace_plts_mean <- get_trace_plots(samp_dt, test_label="mean")
for(plt in trace_plts_mean) plot(plt)
```

```{r}
hist_plts_mean <- get_hist_plot_comparisons(samp_dt, test_label_baseline="exact")
for(plt in hist_plts_mean) plot(plt)
```

```{r}
# Test points subsampled from plug-in mean posterior.
samp_mat <- select_mcmc_samp_mat(samp_dt, test_label="mean", 
                                 param_type="par")[,par_names]
test_info_mean <- get_init_design_list(inv_prob, design_method="subsample", 
                                       N_design=n_design_test,
                                       design_candidates=samp_mat)

em_pred_list_mean <- llik_em$predict_emulator(test_info_mean$input)
```

```{r}
# Plot emulator predictions at points sampled from plug-in mean approximation.
llik_plot_mean <- llik_em$plot_pred_validation(input=test_info_mean$input,
                                                true_llik=test_info_mean$llik,
                                                em_pred_list=em_pred_list_mean, 
                                                interval_method=interval_method,
                                                N_std_dev=N_std_dev, adjustment="none", 
                                                plot_title="Predictions at Plug-in Mean Points") + 
                   geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")


llik_plot_mean_trunc <- llik_em$plot_pred_validation(input=test_info_mean$input,
                                                true_llik=test_info_mean$llik,
                                                em_pred_list=em_pred_list_mean, 
                                                interval_method=interval_method,
                                                N_std_dev=N_std_dev, adjustment="truncated", 
                                                plot_title="Predictions at Plug-in Mean Points (Truncated)") + 
                   geom_hline(yintercept=llik_bounds[2], color="blue", linetype="dashed")

plot(llik_plot_mean)
plot(llik_plot_mean_trunc)
```

## Using Rectified Predictive Distribution
```{r}
mcmc_settings_mean_rect <- list(test_label="mean_rect", approx_type="mean",
                                adjustment="rectified",
                                mcmc_func_name="mcmc_bt_wrapper", 
                                sampler="DEzs", n_itr=50000L, try_parallel=FALSE,  
                                n_chain=4L, defer_ic=TRUE)

# MCMC sampling using exact likelihood: with and without prior truncation.
mcmc_list_mean_rect <- run_mcmc(llik_em, par_prior, mcmc_settings_mean_rect)

# Store samples.
samp_dt_mean_rect <- select_mcmc_itr(mcmc_list_mean_rect$samp, 
                                     itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_mean_rect)
```

```{r}
trace_plts_mean_rect <- get_trace_plots(samp_dt, test_label="mean_rect")
for(plt in trace_plts_mean_rect) plot(plt)
```
```{r}
hist_plts_mean_rect <- get_hist_plot_comparisons(samp_dt, 
                                                 test_label_baseline="exact",
                                                 test_labels="mean_rect")
for(plt in hist_plts_mean_rect) plot(plt)
```

```{r}
hist_plts_mean_compare <- get_hist_plot_comparisons(samp_dt, 
                                                    test_label_baseline="mean",
                                                    test_labels="mean_rect")
for(plt in hist_plts_mean_compare) plot(plt)
```

# Marginal Approximations

## Truncated
```{r}
mcmc_settings_marg_trunc <- list(test_label="marg_trunc", approx_type="marginal",
                                adjustment="truncated",
                                mcmc_func_name="mcmc_bt_wrapper", 
                                sampler="DEzs", n_itr=50000L, try_parallel=FALSE,  
                                n_chain=4L, defer_ic=TRUE)

# MCMC sampling using exact likelihood: with and without prior truncation.
mcmc_list_marg_trunc <- run_mcmc(llik_em, par_prior, mcmc_settings_marg_trunc)

# Store samples.
samp_dt_marg_trunc <- select_mcmc_itr(mcmc_list_marg_trunc$samp, 
                                      itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_marg_trunc)
```

```{r}
trace_plts_marg_trunc <- get_trace_plots(samp_dt, test_label="marg_trunc")
for(plt in trace_plts_marg_trunc) plot(plt)
```
```{r}
trace_plts_marg_trunc_llik <- get_trace_plots(mcmc_list_marg_trunc$info, 
                                              test_label="marg_trunc", 
                                              param_name="llik")
plot(trace_plts_marg_trunc_llik[[1]])
```


```{r}
hist_plts_marg_trunc <- get_hist_plot_comparisons(samp_dt, 
                                                  test_label_baseline="exact",
                                                  test_labels="marg_trunc")
for(plt in hist_plts_marg_trunc) plot(plt)
```


## Rectified
```{r}
mcmc_settings_marg_rect <- list(test_label="marg_rect", approx_type="marginal",
                                adjustment="rectified",
                                mcmc_func_name="mcmc_bt_wrapper", 
                                sampler="DEzs", n_itr=50000L, try_parallel=FALSE,  
                                n_chain=4L, defer_ic=TRUE)

# MCMC sampling using exact likelihood: with and without prior rectation.
mcmc_list_marg_rect <- run_mcmc(llik_em, par_prior, mcmc_settings_marg_rect)

# Store samples.
samp_dt_marg_rect <- select_mcmc_itr(mcmc_list_marg_rect$samp, 
                                      itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_marg_rect)
```

```{r}
trace_plts_marg_rect <- get_trace_plots(samp_dt, test_label="marg_rect")
for(plt in trace_plts_marg_rect) plot(plt)
```

```{r}
hist_plts_marg_rect <- get_hist_plot_comparisons(samp_dt, 
                                                 test_label_baseline="exact",
                                                 test_labels="marg_rect")
for(plt in hist_plts_marg_rect) plot(plt)
```
```{r}

llik_trace_marg_rect <- get_trace_plots(mcmc_list_marg_rect$info,
                                        param_names="llik", 
                                        itr_start=burn_in_mean)

plot(llik_trace_marg_rect[[1]])

```



# Noisy MCMC Algorithms

## MCWMH-Joint, Truncated
```{r}
mcmc_settings_mcwmh_trunc_joint <- list(test_label="mcwmh_trunc_joint", 
                                        mode="mcwmh",
                                        use_joint=TRUE,
                                        adjustment="truncated",
                                        mcmc_func_name="mcmc_noisy_llik", 
                                        n_itr=50000L, try_parallel=FALSE, 
                                        n_chain=4L, accept_rate_target=0.5)

mcmc_list_mcwmh_trunc_joint <- run_mcmc(llik_em, par_prior, 
                                        mcmc_settings_mcwmh_trunc_joint)

# Store samples.
samp_dt_mcwmh_trunc_joint <- select_mcmc_itr(mcmc_list_mcwmh_trunc_joint$samp, 
                                             itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_mcwmh_trunc_joint)
```

```{r}
trace_plts_mcwmh_joint_trunc <- get_trace_plots(samp_dt, 
                                                test_label="mcwmh_trunc_joint")
for(plt in trace_plts_mcwmh_joint_trunc) plot(plt)
```

```{r}
llik_trace_joint_trunc <- get_trace_plots(mcmc_list_mcwmh_trunc_joint$info,
                                          test_labels="mcwmh_trunc_joint",
                                          param_names="llik", 
                                          itr_start=burn_in_mean)
plot(llik_trace_joint_trunc[[1]])
```


```{r}
hist_plts_mcwmh_joint_trunc <- get_hist_plot_comparisons(samp_dt,
                                                         test_label_baseline="exact",
                                                         test_labels="mcwmh_trunc_joint")
for(plt in hist_plts_mcwmh_joint_trunc) plot(plt)
```


```{r}
llik_trace_mcwmh_joint_trunc <- get_trace_plots(mcmc_list_mcwmh_trunc_joint$info, 
                                                test_label="mcwmh_trunc_joint",
                                                param_names="llik")
plot(llik_trace_mcwmh_joint_trunc[[1]])
```

## MCWMH-Joint, Rectified
```{r}
mcmc_settings_mcwmh_rect_joint <- list(test_label="mcwmh_rect_joint", 
                                       mode="mcwmh",
                                       use_joint=TRUE,
                                       adjustment="rectified",
                                       mcmc_func_name="mcmc_noisy_llik", 
                                       n_itr=50000L, try_parallel=FALSE, 
                                       n_chain=4L)

# MCMC sampling using exact likelihood: with and without prior rectation.
mcmc_list_mcwmh_rect_joint <- run_mcmc(llik_em, par_prior, 
                                       mcmc_settings_mcwmh_rect_joint)

# Store samples.
samp_dt_mcwmh_rect_joint <- select_mcmc_itr(mcmc_list_mcwmh_rect_joint$samp, 
                                            itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_mcwmh_rect_joint)
```

```{r}
trace_plts_mcwmh_joint_rect <- get_trace_plots(samp_dt, test_label="mcwmh_rect_joint")
for(plt in trace_plts_mcwmh_joint_rect) plot(plt)
```

```{r}
hist_plts_mcwmh_joint_rec <- get_hist_plot_comparisons(samp_dt, 
                                                       test_label_baseline="exact",
                                                       test_labels="mcwmh_rect_joint")
for(plt in hist_plts_mcwmh_joint_rec) plot(plt)
```


```{r}
llik_trace_joint_rect <- get_trace_plots(mcmc_list_mcwmh_rect_joint$info,
                                         test_labels="mcwmh_rect_joint",
                                         param_names="llik", 
                                         itr_start=burn_in_mean)
plot(llik_trace_joint_rect[[1]])
```

```{r}
test <- get_trace_plots(mcmc_list_mcwmh_rect_joint$info,
                                         test_labels="mcwmh_rect_joint",
                                         param_types="prop", itr_start=25000)
for(plt in test) plot(plt)
```






## MCWMH-Ind, Truncated
```{r}
mcmc_settings_mcwmh_trunc_ind <- list(test_label="mcwmh_trunc_ind", 
                                      mode="mcwmh",
                                      use_joint=FALSE,
                                      adjustment="truncated",
                                      mcmc_func_name="mcmc_noisy_llik", 
                                      n_itr=50000L, try_parallel=FALSE, 
                                      n_chain=4L)

# MCMC sampling using exact likelihood: with and without prior rectation.
mcmc_list_mcwmh_trunc_ind <- run_mcmc(llik_em, par_prior, 
                                      mcmc_settings_mcwmh_trunc_ind)

# Store samples.
samp_dt_mcwmh_trunc_ind <- select_mcmc_itr(mcmc_list_mcwmh_trunc_ind$samp, 
                                           itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_mcwmh_trunc_ind)
```

```{r}
trace_plts_mcwmh_ind_trunc <- get_trace_plots(samp_dt, test_label="mcwmh_trunc_ind")
for(plt in trace_plts_mcwmh_ind_trunc) plot(plt)
```

```{r}
hist_plts_mcwmh_ind_trunc <- get_hist_plot_comparisons(samp_dt, 
                                                       test_label_baseline="exact",
                                                       test_labels="mcwmh_trunc_ind")
for(plt in hist_plts_mcwmh_ind_trunc) plot(plt)
```


## MCWMH-Ind, Rectified
```{r}
mcmc_settings_mcwmh_rect_ind <- list(test_label="mcwmh_rect_ind", 
                                      mode="mcwmh",
                                      use_joint=FALSE,
                                      adjustment="rectified",
                                      mcmc_func_name="mcmc_noisy_llik", 
                                      n_itr=50000L, try_parallel=FALSE, 
                                      n_chain=4L)

# MCMC sampling using exact likelihood: with and without prior rectation.
mcmc_list_mcwmh_rect_ind <- run_mcmc(llik_em, par_prior, 
                                     mcmc_settings_mcwmh_rect_ind)

# Store samples.
samp_dt_mcwmh_rect_ind <- select_mcmc_itr(mcmc_list_mcwmh_rect_ind$samp, 
                                          itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_mcwmh_rect_ind)
```

```{r}
trace_plts_mcwmh_ind_rect <- get_trace_plots(samp_dt, test_label="mcwmh_rect_ind")
for(plt in trace_plts_mcwmh_ind_rect) plot(plt)
```

```{r}
hist_plts_mcwmh_ind_rect <- get_hist_plot_comparisons(samp_dt, 
                                                      test_label_baseline="exact",
                                                      test_labels="mcwmh_rect_ind",
                                                      chain_idcs=4)
for(plt in hist_plts_mcwmh_ind_rect) plot(plt)
```

```{r}
llik_trace_ind_rect <- get_trace_plots(mcmc_list_mcwmh_rect_ind$info,
                                       test_labels="mcwmh_rect_ind",
                                       param_names="llik")
plot(llik_trace_ind_rect[[1]])
```


## MCWMH-Joint, No truncation.
```{r}
mcmc_settings_mcwmh_joint <- list(test_label="mcwmh_joint", 
                                  mode="mcwmh",
                                  use_joint=TRUE,
                                  adjustment="none",
                                  mcmc_func_name="mcmc_noisy_llik", 
                                  n_itr=50000L, try_parallel=FALSE, 
                                  n_chain=4L)

mcmc_list_mcwmh_joint <- run_mcmc(llik_em, par_prior, 
                                  mcmc_settings_mcwmh_joint)

# Store samples.
samp_dt_mcwmh_joint <- select_mcmc_itr(mcmc_list_mcwmh_joint$samp, 
                                       itr_start=burn_in_mean)
samp_dt <- combine_samp_dt(samp_dt, samp_dt_mcwmh_joint)
```

```{r}
trace_plts_mcwmh_joint <- get_trace_plots(samp_dt, test_label="mcwmh_joint",
                                          chain_idcs=2)
for(plt in trace_plts_mcwmh_joint) plot(plt)
```

```{r}
hist_plts_mcwmh_joint <- get_hist_plot_comparisons(samp_dt, 
                                                   test_label_baseline="exact",
                                                   test_labels="mcwmh_joint")
for(plt in hist_plts_mcwmh_joint) plot(plt)
```

```{r}
llik_trace_joint <- get_trace_plots(mcmc_list_mcwmh_joint$info,
                                    test_labels="mcwmh_joint",
                                    param_names="llik")
plot(llik_trace_joint[[1]])
```

Things to look into for noisy MCMC performance: 
- Investigate the trajectory of the acceptance rate.
- Consider changing the adaptation settings (target acceptance rate, etc.)
- Look into GP pred mean and variance at the noisy MCMC posterior samples; do the algs tend to sample from regions with high mean? Variance?
- Return to ratio analysis to better understand the algorithms.
- Try running noisy MCMC for the 1d vis example








