---
title: "JPL UQ Talk"
author: "Andrew Roberts"
date: '2024-09-10'
output: html_document
---

This RMarkdown generates plots for my talk "UQ for Remote Sensing Breakout" on 
09/17/2024. For this 20 minute presentation, I use two examples: the 1d test 
that serves as the visual 1d example in the GP emulator paper; the plots for 
this example are taken from 
*gp-calibration/scripts/gp_post_approx_paper/test_1d_visualization.Rmd* (they 
are not generated here). The second example features a 3-dimensional input 
space; this example is run utilizing the script *run_vsem_example.r*, and the 
results are read into this document from the directory 
*gp-calibration/output/gp_post_approx_paper/vsem/test_d3_p1_N30_LHS*. 

```{r, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(lhs)
library(ggplot2)
library(viridis)
library(parallel)
library(gridExtra)
library(data.table)

experiment_tag <- "vsem"
run_tag <- "test_d3_p1_N30_LHS"
base_dir <- file.path("/projectnb", "dietzelab", "arober", "gp-calibration")
src_dir <- file.path(base_dir, "src")
output_dir <- file.path(base_dir, "output", "gp_post_approx_paper", 
                        experiment_tag, run_tag)
jpl_talk_dir <- file.path(base_dir, "output", "jpl_talk_2024")

source(file.path(src_dir, "gpWrapper.r"))
source(file.path(src_dir, "general_helper_functions.r"))
source(file.path(src_dir, "gp_helper_functions.r"))
source(file.path(src_dir, "inv_prob_test_functions.r"))
source(file.path(src_dir, "statistical_helper_functions.r"))
source(file.path(src_dir, "plotting_helper_functions.r"))
source(file.path(src_dir, "mcmc_helper_functions.r"))
source(file.path(src_dir, "llikEmulator.r"))
source(file.path(src_dir, "gp_emulator_functions.r"))
source(file.path(src_dir, "sim_study_functions.r"))
source(file.path(src_dir, "mcmc_calibration_functions.r"))
source(file.path(src_dir, "seq_design_gp.r"))
source(file.path(src_dir, "seq_design.r"))
source(file.path(src_dir, "gp_mcmc_functions.r"))

# Plot settings applied to all plots. 
design_pt_size <- 4
line_thickness <- 1.5
design_color <- "red"
xlim <- c(-10, 10)
ylim <- c(0.0777, 0.4)
approx_type_labels <- c("mcwmh-ind", "mcwmh-joint", "gp-mean", "gp-marg")
approx_type_colors <- c("blue", "red", "orange", "purple")
approx_types <- c("mcwmh-ind-em_llik_const_GaussQuad", "mcwmh-joint-em_llik_const_GaussQuad",
                  "gp-mean-em_llik_const_GaussQuad", "gp-marg-em_llik_const_GaussQuad")
approx_type_labels <- setNames(approx_type_labels, approx_types)
approx_type_colors <- setNames(approx_type_colors, approx_types)


# Settings for MCMC plotting. 
burn_in_start <- 100000L

# Settings for confidence intervals displayed in plots. 
interval_method <- "pm_std_dev"
N_std_dev <- 1.64

# I only display plots for one of the log-likelihood emulators in the slides 
# for the talk. This variable determines which emulator is used.
em_name_talk <- "em_llik_const_GaussQuad"
```

# Summary of Inverse Problem 

```{r}

#
# Plot model outputs and observations.  
#

dir.create(jpl_talk_dir)

load(file.path(output_dir, "inv_prob_list.RData"))

df_inv_prob_data <- data.frame(t=inv_prob$times, y_true=drop(inv_prob$output_true), 
                               y=drop(inv_prob$y))
plt_outputs <- ggplot(data=df_inv_prob_data) +
                      geom_point(aes(x=t, y=y), color="red") + 
                      geom_line(aes(x=t, y=y_true), color="black") + 
                      ylab("output") + ggtitle("True and Observed y")

plot(plt_outputs)
```

```{r}
# Print settings. 

load(file.path(output_dir, "design_info.RData"))

print("Settings:")
print(paste0("Parameter space dimension: ", nrow(inv_prob$par_prior)))
print(paste0("Number design points: ", design_info$N_design))
print(paste0("Design method: ", design_info$design_method))

print("Priors and true parameter:")
inv_prob$par_prior$par_true <- inv_prob$par_true
print(inv_prob$par_prior)
```
# Inverse Problem Solution via Exact MCMC (no emulation)

```{r}
#
# MCMC Plots: exact.  
#

samp_dt <- fread(file.path(output_dir, "mcmc_samp.csv"))

# MCMC Plots: exact. 
trace_plots_exact <- get_trace_plots(samp_dt, test_label="exact", 
                                     param_types="par", burn_in_start=burn_in_start)
hist_plots_exact <- get_hist_plot_comparisons(samp_dt, test_labels="exact", 
                                              param_types="par", burn_in_start=burn_in_start)
corr_plots_exact <- get_mcmc_2d_density_plots(samp_dt, test_labels="exact", 
                                              burn_in_start=burn_in_start)$exact

for(plt in trace_plots_exact) plot(plt) 
for(plt in hist_plots_exact) plot(plt)
for(plt in corr_plots_exact) plot(plt)

```

# Emulator Summary and Validation

```{r}
#
# Analyzing Log-likelihood emulator performance. 
#

load(file.path(output_dir, "llik_em_list.RData"))
load(file.path(output_dir, "emulator_pred_list.RData"))
load(file.path(output_dir, "test_info.RData"))
em_names <- names(llik_em_list)

for(em_name in em_names) {
  print(em_name)
  cat(llik_em_list[[em_name]]$emulator_model$summarize(), "\n")
}
```

```{r}
# Log-likelihood emulator predictions at test points. 

for(em_name in em_names) {
  plt <- llik_em_list[[em_name]]$plot_pred_validation(test_info$input, true_llik=test_info$llik, 
                                                      emulator_pred_list=emulator_pred_list[[em_name]], 
                                                      plot_type="llik", include_interval=TRUE, 
                                                      interval_method=interval_method, N_std_dev=N_std_dev, 
                                                      plot_title=paste0("llik predictions: ", em_name))
  plot(plt)
}
```

```{r}
# Plot 1d projections: log-likelihood emulator distribution  

llik_func <- function(u) llik_exact$assemble_llik(u)

N_test_proj <- 100L
input_list_proj <- list()
for(par_name in inv_prob$par_names) {
  prior_bounds <- c(inv_prob$par_prior[par_name,"param1"], 
                    inv_prob$par_prior[par_name,"param2"])
  input_list_proj[[par_name]] <- seq(prior_bounds[1], prior_bounds[2], length.out=N_test_proj)
}

# Plot 1d projections. 
for(em_name in names(llik_em_list)) {

  plts <- llik_em_list[[em_name]]$plot_1d_projection(input_list_proj=input_list_proj,
                                                     include_design=TRUE, plot_title=em_name,
                                                     interval_method=interval_method, 
                                                     N_std_dev=N_std_dev, 
                                                     llik_func_true=llik_func)
  for(plt in plts) plot(plt)
}
```

```{r}
# Plot 1d projection onto one variable for slides.
proj_var <- inv_prob$par_names

plt_proj_talk <- llik_em_list[[em_name_talk]]$plot_1d_projection(proj_var, input_list_proj=input_list_proj,
                                                                 include_design=FALSE, plot_title=em_name_talk,
                                                                 interval_method=interval_method, 
                                                                 N_std_dev=N_std_dev,
                                                                 llik_func_true=llik_func) 
                                                                 
plt_proj_talk_mean_approx <- llik_em_list[[em_name_talk]]$plot_1d_projection(proj_var,
                                                                             input_list_proj=input_list_proj, 
                                                                             plot_type="lik_approx", 
                                                                             approx_types="mean",
                                                                             include_design=TRUE, 
                                                                             llik_func_true=llik_func,
                                                                             plot_title=em_name_talk)
                                                                             
plt_proj_talk_marg_approx <- llik_em_list[[em_name_talk]]$plot_1d_projection(proj_var, input_list_proj=input_list_proj, 
                                                                             plot_type="lik_approx", 
                                                                             approx_types="marginal",
                                                                             include_design=TRUE,
                                                                             llik_func_true=llik_func,
                                                                             plot_title=em_name_talk)

lbl_mean <- "gp-mean-em_llik_const_GaussQuad"
lbl_marg <- "gp-marg-em_llik_const_GaussQuad"
for(i in seq_along(proj_var)) {
  # Format plots. 
  plt <- ggformat_journal(plt_proj_talk[[i]], legend_position="none")
  plot_mean_approx <- ggformat_journal(plt_proj_talk_mean_approx[[i]], legend_position="bottom") + 
                      scale_color_manual(labels=c("mean"="gp-mean"), 
                                         values=c("mean"=unname(approx_type_colors[lbl_mean])))
  plot_marg_approx <- ggformat_journal(plt_proj_talk_marg_approx[[i]], legend_position="bottom") + 
                      scale_color_manual(labels=c("marginal"="gp-marg"), 
                                         values=c("marginal"=unname(approx_type_colors[lbl_marg])))
  
  # Display plots.
  plot(plt)
  plot(plot_mean_approx)
  plot(plot_marg_approx)
  
  # Save plots. 
  base_name <- paste("proj1d", proj_var[i], em_name_talk, sep="_")
  ggsave(file.path(jpl_talk_dir, paste0(base_name, "llikdist.png")), plt)
  ggsave(file.path(jpl_talk_dir, paste0(base_name, "meanapprox.png")), plot_mean_approx)
  ggsave(file.path(jpl_talk_dir, paste0(base_name, "margapprox.png")), plot_marg_approx)
}


```

```{r}
# Plot 1d projections: deterministic likelihood approximations. 

approx_types <- c("mean", "marginal")

for(em_name in names(llik_em_list)) {
  plts <- llik_em_list[[em_name]]$plot_1d_projection(input_list_proj=input_list_proj, 
                                                     plot_type="lik_approx", 
                                                     approx_types=approx_types,
                                                     include_design=TRUE, plot_title=em_name)
  for(plt in plts) plot(plt)
}
```

# GP-Induced Approximate Posterior Samples

```{r}
#
# MCMC Trace Plots: approximate. 
#

test_labels_approx <- setdiff(unique(samp_dt$test_label), "exact")
trace_plots_approx <- get_trace_plots(samp_dt, test_label=test_labels_approx,
                                      param_types="par", burn_in_start=burn_in_start)

for(plt in trace_plots_approx) plot(plt) 
```

```{r}
# Histograms of marginal samples. 
test_plts <- get_hist_plot_comparisons(samp_dt, burn_in_start=burn_in_start, 
                                       param_types="par", test_labels=test_labels_approx, 
                                       param_names="KEXT", test_label_baseline="exact")

for(plt in test_plts) plot(plt)
```


```{r}
# 1d KDE plots. 

kde_plts <- get_1d_kde_plot_comparisons(samp_dt, burn_in_start=burn_in_start, 
                                        param_types="par",
                                        test_label_baseline="exact",
                                        test_labels=test_labels_approx,
                                        xlab="parameter", ylab="kde", save_dir=NULL) 

for(plt in kde_plts) plot(plt)
```

```{r}
# Save KDE plots for talk.

# Noisy MCMC algorithms. 
lbl_noisy <- c("mcwmh-ind-em_llik_const_GaussQuad", "mcwmh-joint-em_llik_const_GaussQuad")
for(par_name in inv_prob$par_names) {
  plt <- get_1d_kde_plot_comparisons(samp_dt, burn_in_start=burn_in_start, 
                                     param_types="par", param_names=par_name,
                                     test_label_baseline="exact",
                                     test_labels=lbl_noisy,
                                     xlab="parameter", ylab="kde")[[1]] + xlab(par_name) + 
                                     scale_color_manual(labels=approx_type_labels[lbl_noisy],
                                                        values=approx_type_colors[lbl_noisy])
                                                        
  plt <- ggformat_journal(plt, legend_position="bottom")
  plot(plt)
  ggsave(file.path(jpl_talk_dir, paste0("mcwmh_GaussQuad_kde_", par_name, ".png")), plt)
}


# Mean approximations.
lbl_mean <- "gp-mean-em_llik_const_GaussQuad"
for(par_name in inv_prob$par_names) {
  plt <- get_1d_kde_plot_comparisons(samp_dt, burn_in_start=burn_in_start, 
                                     param_types="par", param_names=par_name,
                                     test_label_baseline="exact",
                                     test_labels=lbl_mean,
                                     xlab="parameter", ylab="kde")[[1]] + xlab(par_name) + 
                                     scale_color_manual(labels=approx_type_labels[lbl_mean],
                                                        values=approx_type_colors[lbl_mean])
  
  plt <- ggformat_journal(plt, legend_position="bottom")
  plot(plt)                                             
  ggsave(file.path(jpl_talk_dir, paste0("gp-mean_GaussQuad_kde_", par_name, ".png")), plt)
}

# Marginal approximations.
lbl_marg <- "gp-marg-em_llik_const_GaussQuad"
for(par_name in inv_prob$par_names) {
  plt <- get_1d_kde_plot_comparisons(samp_dt, burn_in_start=burn_in_start, 
                                     param_types="par", param_names=par_name,
                                     test_label_baseline="exact",
                                     test_labels=lbl_marg,
                                     ylab="kde")[[1]] + xlab(par_name) + 
                                     scale_color_manual(labels=approx_type_labels[lbl_marg],
                                                        values=approx_type_colors[lbl_marg])
                                               
  plt <- ggformat_journal(plt, legend_position="bottom")
  plot(plt)                                             
  ggsave(file.path(jpl_talk_dir, paste0("gp-marg_GaussQuad_kde_", par_name, ".png")), plt)
}

```



```{r}
# mcmc_par_stats <- compute_mcmc_param_stats(samp_dt, burn_in_start=burn_in_start, 
#                                            test_labels=NULL, param_types="par")
#                          
# mcmc_errs_list <- compute_mcmc_comparison_metrics(samp_dt, "exact", "marg-fwd", c("mean", "cov"), 
#                                                   burn_in_start=burn_in_start, param_types="par")
                                                  

```


```{r}
#
# 1d marginal plots (KDEs)
#

kde_plts <- get_1d_kde_plot_comparisons(samp_dt, N_kde_pts=100, burn_in_start=burn_in_start, 
                                        param_types="par", param_names=NULL,
                                        test_label_baseline="exact",
                                        xlab="parameter", ylab="kde", save_dir=NULL) 

for(plt in kde_plts) plot(plt)
```

```{r}
scatter_stat_plts <- get_mcmc_moments_scatter_plot_comparisons(samp_dt, test_label_baseline="exact",
                                                               burn_in_start=burn_in_start, param_types="par")

for(plt in scatter_stat_plts) plot(plt)
```

```{r}
tst_lbls <- approx_types
coverage_plt_list <- get_1d_coverage_plots(samp_dt, "exact", test_labels=tst_lbls, 
                                           burn_in_start=burn_in_start,
                                           param_types="par", probs=seq(.5, .99, .05))
coverage_plts <- coverage_plt_list$plots

for(i in seq_along(coverage_plts)) {
  plot_name <- names(coverage_plts)[i]
  plt <- ggformat_journal(coverage_plts[[i]], remove_title=FALSE, legend_position="bottom") + 
         scale_color_manual(labels=approx_type_labels[tst_lbls],
                            values=approx_type_colors[tst_lbls])
  plot(plt)
  ggsave(file.path(jpl_talk_dir, paste0("coverage_", plot_name, ".png")), plt)
}
```









