---
title: "Emulator Metrics Illustration"
author: "Andrew Roberts"
date: '2023-06-07'
output: html_document
---

```{r, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lhs)
library(hetGP)
library(mlegp)
library(ggplot2)
library(viridis)
library(gridExtra)
library(data.table)
library(BayesianTools)

source("numerical_experiment_functions.r")
source("mcmc_calibration_functions.r")
source("gp_emulator_functions.r")
```


# Computer Model and Data 

```{r}
# Synthetic data generation
computer_model_data <- generate_vsem_test_case(4)
```

```{r}
print(computer_model_data$ref_pars[computer_model_data$pars_cal_sel,])
```

```{r echo = FALSE}
for(output_var in computer_model_data$output_vars) {
 plotTimeSeries(observed = computer_model_data$data_obs[, output_var],
                predicted = computer_model_data$data_ref[, output_var], main = output_var) 
}
```

# Priors 

## Calibration Parameters 
```{r, echo = FALSE} 
# Priors 
theta_prior_params <- computer_model_data$ref_pars[computer_model_data$pars_cal_sel,]
theta_prior_params[, "dist"] <- c("Uniform", "Uniform")
theta_prior_params[,"param1"] <- theta_prior_params[,"lower"]
theta_prior_params[,"param2"] <- theta_prior_params[,"upper"]
theta_prior_params <- theta_prior_params[, c("dist", "param1", "param2")]

print(theta_prior_params)
```


## Likelihood Parameters 
```{r}
sig2_prior_info <- get_IG_priors_numerical_test(sig2_true = diag(computer_model_data$Sig_eps), 
                                                bias_frac = c(0.1, -0.15), bins = 50,
                                                coef_var = c(0.3, 0.5), return_prior_plots = TRUE, 
                                                output_variables = computer_model_data$output_vars)
sig_eps_prior_params <- sig2_prior_info$prior
plts <- sig2_prior_info$plots

grid.arrange(arrangeGrob(grobs = plts, nrow = 1))
```

# Exact MCMC Samples.

## Exact MCMC. 
```{r}
N_mcmc_exact <- 50000

time_start <- proc.time()
samp_mcmc_exact <- mcmc_calibrate_product_lik(computer_model_data = computer_model_data, 
                                              theta_prior_params = theta_prior_params, 
                                              learn_sig_eps = TRUE,
                                              sig_eps_prior_params = sig_eps_prior_params,
                                              N_mcmc = N_mcmc_exact)
mcmc_exact_runtime <- (proc.time() - time_start)[["elapsed"]]

samp_exact_df <- as.data.frame(samp_mcmc_exact$samp)
colnames(samp_exact_df) <- paste("exact", colnames(samp_exact_df), sep = "_")

print(paste0("Exact MCMC runtime: ", mcmc_exact_runtime, " seconds."))
```

```{r}
burn_in_exact <- c(exact = 10000)

mcmc_exact_trace_plts <- get_trace_plots(samp_exact_df, burn_in_start = burn_in_exact, param_types = "theta")
grid.arrange(arrangeGrob(grobs = mcmc_exact_trace_plts, nrow = 1))
```

```{r}
mcmc_exact_trace_plts_sig_eps <- get_trace_plots(samp_exact_df, burn_in_start = burn_in_exact, param_types = "sig_eps")
grid.arrange(arrangeGrob(grobs = mcmc_exact_trace_plts_sig_eps, nrow = 1))
```
      
# True Response Surfaces. 

```{r}
# Evaluate over dense grid, and overlay posterior contours. 
theta_prior_design <- get_input_output_design(N_points = 51^2, 
                                              computer_model_data = computer_model_data, 
                                              theta_prior_params = theta_prior_params, 
                                              design_method = "grid")
```

```{r}
samp_exact_theta <- as.matrix(samp_exact_df[burn_in_exact:N_mcmc_exact, c("exact_theta_LAR", "exact_theta_KEXT")])
colnames(samp_exact_theta) <- c("LAR", "KEXT")
```


### Prior Plot. 
```{r}
plt <- get_2d_response_surface_plot(computer_model_data - computer_model_data, 
                                    theta_vals = theta_prior_design$inputs, 
                                    param_names = computer_model_data$pars_cal_names, 
                                    response_surface = "prior", 
                                    raster = TRUE, 
                                    theta_prior_params = theta_prior_params,
                                    point_coords = computer_model_data$theta_true, 
                                    samples_kde = samp_exact_theta, 
                                    scale_inputs = TRUE, 
                                    input_bounds = theta_prior_design$input_bounds)[[1]]

plot(plt)

```


### SSR Plot. 
```{r}
plts <- get_2d_response_surface_plot(computer_model_data - computer_model_data, 
                                     theta_vals = theta_prior_design$inputs, 
                                     param_names = computer_model_data$pars_cal_names, 
                                     response_surface = "SSR", 
                                     SSR_vals = theta_prior_design$outputs, 
                                     raster = TRUE, 
                                     point_coords = computer_model_data$theta_true, 
                                     samples_kde = samp_exact_theta, 
                                     scale_inputs = TRUE, 
                                     input_bounds = theta_prior_design$input_bounds)

grid.arrange(arrangeGrob(grobs=plts, nrow=1))

```

### Likelihood Plot. 
```{r}
plt <- get_2d_response_surface_plot(computer_model_data = computer_model_data, 
                                    theta_vals = theta_prior_design$inputs, 
                                    param_names = computer_model_data$pars_cal_names, 
                                    response_surface = "likelihood", 
                                    SSR_vals = theta_prior_design$outputs, 
                                    combine_outputs = TRUE, 
                                    raster = TRUE, 
                                    point_coords = computer_model_data$theta_true, 
                                    samples_kde = samp_exact_theta, 
                                    scale_inputs = TRUE, 
                                    input_bounds = theta_prior_design$input_bounds, 
                                    theta_prior_params = theta_prior_params)[[1]]

plot(plt)
```

### Individual likelihoods. 
```{r}
plts <- get_2d_response_surface_plot(computer_model_data = computer_model_data, 
                                    theta_vals = theta_prior_design$inputs, 
                                    param_names = computer_model_data$pars_cal_names, 
                                    response_surface = "likelihood", 
                                    SSR_vals = theta_prior_design$outputs, 
                                    combine_outputs = FALSE, 
                                    raster = TRUE, 
                                    point_coords = computer_model_data$theta_true, 
                                    samples_kde = samp_exact_theta, 
                                    scale_inputs = TRUE, 
                                    input_bounds = theta_prior_design$input_bounds, 
                                    theta_prior_params = theta_prior_params)

grid.arrange(arrangeGrob(grobs=plts, nrow=1))
```


### Posterior Plot. 
```{r}
plt <- get_2d_response_surface_plot(computer_model_data = computer_model_data, 
                                    theta_vals = theta_prior_design$inputs, 
                                    param_names = computer_model_data$pars_cal_names, 
                                    response_surface = "posterior", 
                                    SSR_vals = theta_prior_design$outputs, 
                                    raster = TRUE, 
                                    point_coords = computer_model_data$theta_true, 
                                    samples_kde = samp_exact_theta, 
                                    scale_inputs = TRUE, 
                                    input_bounds = theta_prior_design$input_bounds, 
                                    theta_prior_params = theta_prior_params)

plot(plt)

```

# Emulator: Varying number of design points. 

```{r, echo = FALSE}
# Emulator settings
emulator_settings <- data.frame(gp_lib = c("hetGP"), 
                                kernel = "Gaussian", 
                                transformation_method = c("truncated"),
                                scale_X = TRUE, 
                                normalize_y = TRUE)
print(emulator_settings)
```

We consider three Latin hypercube designs, with 20, 30, and 40 design points, respectively. We consider 50 replications of each design so that the error metrics can be averaged over the replicates.  
```{r}

# Number of design replications. 
design_reps <- 50

# The three design specifications to be used. 
design_settings <- data.frame(N_design = c(20, 30, 30), 
                              design_method = c("LHS", "LHS", "sample_candidates"), 
                              scale_inputs = TRUE, 
                              normalize_response = TRUE)
                              

# Get designs. 
design_list <- get_design_list(design_settings = design_settings, 
                               computer_model_data = computer_model_data, 
                               theta_prior_params = theta_prior_params,
                               reps = design_reps, 
                               design_candidates = samp_exact_theta)

```

```{r}
emulator_info_list<- fit_emulator_design_list(emulator_settings[1,], design_list)
```

```{r}
# Validation sets sampled from prior. 
N_test_points <- 20
N_test_sets <- 100

transformation_method <- ifelse(any(emulator_settings == "LNP"), "LNP", NA_character_)
validation_list <- get_design_list_test_data(N_test_points = N_test_points, 
                                             N_test_sets = N_test_sets, 
                                             design_method = "LHS", 
                                             design_list = design_list, 
                                             computer_model_data = computer_model_data, 
                                             theta_prior_params = theta_prior_params, 
                                             transformation_method = transformation_method)

```

```{r}
# Validation sets sampled from posterior. 
validation_list_post <- get_design_list_test_data(N_test_points = N_test_points, 
                                                  N_test_sets = N_test_sets, 
                                                  design_method = "sample_candidates", 
                                                  design_list = design_list, 
                                                  computer_model_data = computer_model_data, 
                                                  theta_prior_params = theta_prior_params, 
                                                  transformation_method = transformation_method, 
                                                  design_candidates = samp_exact_theta)
```

```{r}
emulator_metrics <- get_emulator_comparison_metrics_validation_list(emulator_info_list = emulator_info_list, 
                                                                   X_test_list = lapply(validation_list, function(l) l$inputs),
                                                                   Y_test_list = lapply(validation_list, function(l) l$outputs),
                                                                   metrics = c("rmse", "srmse", "nlpd_pointwise", "crps"), 
                                                                   scale_inputs = TRUE, 
                                                                   output_variables = computer_model_data$output_vars, 
                                                                   include_nug = TRUE)
```

```{r}
emulator_metrics_post <- get_emulator_comparison_metrics_validation_list(emulator_info_list = emulator_info_list, 
                                                                         X_test_list = lapply(validation_list_post, function(l) as.matrix(l$inputs)),
                                                                         Y_test_list = lapply(validation_list_post, function(l) as.matrix(l$outputs)),
                                                                         metrics = c("rmse", "srmse", "nlpd_pointwise", "crps"), 
                                                                         scale_inputs = TRUE, 
                                                                         output_variables = computer_model_data$output_vars, 
                                                                         include_nug = TRUE)
```

We consider sampling validation points based on the prior and posterior. Validation points are sampled so that they satisfy the bounds 
for all of the designs that are being tested (i.e. no emulators will be extrapolating). Since we sampled one of the designs from the 
posterior (which is highly concentrated) this will have the effect that the samples based on the prior below are actually much more 
similar to the posterior samples. 

### Validation points based on prior. 
```{r}
plts <- plot_emulator_comparison_violin_plots(emulator_metrics, fill_col = "design",
                                              metrics = c("crps", "rmse", "nlpd_pointwise"), 
                                              output_variables = computer_model_data$output_vars, 
                                              include_points = FALSE, 
                                              nrow = 2)
grid.arrange(plts)
```

### Validaton points sampled from posterior. 
```{r}
plts <- plot_emulator_comparison_violin_plots(emulator_metrics_post, fill_col = "design",
                                              metrics = c("crps", "rmse", "nlpd_pointwise"), 
                                              output_variables = computer_model_data$output_vars, 
                                              include_points = FALSE, 
                                              nrow = 2)
grid.arrange(plts)
```





